---
title: "Lab 3 Activity"
output: pdf_document
header-includes:
    - \usepackage{setspace}
---



```{r include=FALSE}
library(tidyverse)
```

For this activity we will be looking at some data about [the 1907 Romanian peasant revolt](https://en.wikipedia.org/wiki/1907_Romanian_peasants%27_revolt#:~:text=A%20peasant%20revolt%20(Romanian%3A%20Răscoala,including%20as%20far%20as%20Oltenia.){target="_blank"}! The data is called `Chirot` and is stored in the `carData` package. Run `help("Chirot")` to find out more about the data.


**1.** Load the `carData` and create a separate `data.frame` object called `rev_int` that only contains the the `intensity` and `inequality` variables from the `chirot` data.

```{r include=FALSE}
library(carData)

# MANY ways of doing this

rev_int <- Chirot[,c("intensity","inequality")]

# or also any of tyhe lines below works

# rev_int <- Chirot[,c(1,5)]
# rev_int <- dplyr::select(.data = Chirot, intensity, inequality)
# rev_int <- data.frame(Chirot$intensity, Chirot$inequality)
# rev_int <- cbind(Chirot$intensity, Chirot$inequality)

```

**2.** Find the correaltion between `inequality` and `intesity`. Is it significantly different from 0? What is the 95% confidence interval?

```{r include=FALSE}

cor <- psych::corr.test(rev_int)

# correlation
cor$r[2]

# significance and 95%CI
cor$ci


# not significant
```



**3.**



**4.**




**5.**



**6.**


\newpage


## Some R Practice












