---
title: "Lab 2 Activity"
output: pdf_document
header-includes:
    - \usepackage{setspace}
---


```{r include=FALSE}
library(tidyverse)
```


In this lab activity we are going to use the `attitude` dataset, which is already loaded into R. 

**1.** Run `help("attitude")` to get some info on the data and the meaning of each variable. You can also run `View(attitude)` to open the data in the data viewer window.

```{r include=FALSE}
help("attitude")

dat <- attitude
```


**1.** Run `help("attitude")` to get some info on the data and the meaning of each variable. You can also run `View(attitude)` to open the data in the data viewer window.


**2.** plot `rating` on the $y$-axis and `complaints` on the $x$-axis. If you were to run a regression, do you think the slope ($b_1$) will be positive or negative? Why? 

```{r include=FALSE}

ggplot(attitude,  
 aes(x = complaints, y = rating)) +
 geom_point() +
 theme_classic()


# positive as when complaint handling is more positive, ratings increase 
```

**3.** Run a linear regression with `complaints` predicting `rating`. 

```{r include=FALSE}
reg <- lm(rating ~ complaints, attitude_std)

summary(reg)
```


- What do you conclude about the relation between rating and complaints? 

```{r include=FALSE}
# there is a significantly positive relation between positive handling of complaints and rating
```

- How much would we expect `rating` increase to be on average if `complaints` increased by 3 units?

```{r include=FALSE}
# rating would increase by

coef(reg)[2]*3
```

**4.** Run a *standardized* regression with `complaints` predicting `rating`. What changes do you see in the `summary()` output?

```{r include=FALSE}
attitude_std <- data.frame(scale(attitude))

reg_std <- lm(rating ~ complaints, attitude_std)

summary(reg_std)

# intercept is now 0 and the slope is now .83. Everything else remains the same.
```

**5.** What is the predicted value of `rating` in standardized units when `complaints` is 1 standard deviation below average?

```{r include=FALSE}
# we would expect ratings to be
coef(reg_std)[2]*-1
```

**Tricky questions**

- can you convert this value of `rating` back into unstandardized units? (HINT: you will need to use the mean and standard deviation of the original `rating` variable)

```{r}

mean_rtg <- mean(attitude$rating) 
sd_rtg <- sd(attitude$rating)


# the formula to do so is Mean + (SD_rating*predicted_y_value_std)

mean_rtg + (sd_rtg * (coef(reg_std)[2]*-1))


# this is because -.83 stands for "-.83 standard deviations below the mean of rating", which is   54.56 in unstandardized units.

# for IQ it would be 100 + (15*-.83) = 87.55

```

- How do you get the same value using the unstandardized regression equation? (HINT: you need to use the mean value and standard deviation of the `complaints` variable)

```{r}
mean_cmpl <- mean(attitude$complaints) 
sd_cmpl <- sd(attitude$complaints)

# simpler here because we just need to multiply by -1 because we are "one standard deviation below the average complaint handling" 

X_value <- mean_cmpl + (sd_cmpl * -1)

coef(reg)[1] + coef(reg)[2]*X_value
```


\newpage


## Some R Practice












