---
title: "Lab 4 Activity"
output: pdf_document
header-includes:
    - \usepackage{setspace}
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```

\onehalfspacing

```{r include=FALSE}
library(tidyverse)
```

Let's look at the `attitude` dataset, which we used in the activity for Lab 2. This dataset is always loaded into R.

**1.** Run `help("attitude")` to get some info on the data and the meaning of each variable. You can also run `View(attitude)` to open the data in the data viewer window. Then, create a dataset called `dat` that only includes the `rating`, `privileges`, and `complaints` variables.

```{r include=FALSE}
help("attitude")

dat <- attitude[, c("rating", "privileges", "complaints")]
```

**2** Use the `ggpairs()` function from `GGally` to visualize the relation and distribution of the 3 variables in the the `dat` object. Do these variables seem linearly related? If so, do they seem positively or negatively related?

```{r}
GGally::ggpairs(dat)

# yes very related
```

**4.** Run two individual regressions, with one regression having the variable `privileges` predicting `rating` and the other having the `complaints` variable predicting `rating`. What are the values of the two regression slopes? Do they match your expectations? Are the values of the slopes significant?

```{r}
reg_priv  <- lm(rating ~ privileges, data = dat) 

reg_compl <- lm(rating ~ complaints, data = dat)   


car::S(reg_priv)
car::S(reg_compl)

# both slopes are positive and significant. This is in line with expectations given that the previous plots both showed postive relations among the variables. 
```

**5.** Run a multiple linear regression where both `privileges` and `complaints` predict `rating`. What are the values of the slopes now? Is there anything that you find surprising?

```{r}
reg_full <- lm(rating ~ complaints + privileges, data = dat)

car::S(reg_full)

# only the slope of complaints is significant now. Further, the slope of privileges is now negative (!). This happens because complaints is positively correlated with privileges. 
```

**6.** Visualize your regression model with an interactive 3D plot that includes a regression plane. Note that this plot will not be visible in a PDF once you knit (why? PDFs are not interactive). If you are wondering why the slope of `privileges` is now negative, looking at the plot from a specific angle may give you some insight.

```{r}
library(FabioFun)
# 
# nice_3D_plot(y =  dat$rating,
#              x1 = dat$complaints,
#              x2 = dat$privileges,
#              axis_names = c("Rating", 
#                             "Complaints", 
#                             "Privileges"),
#              reg_plane = TRUE)
```

-   What are some of the points with the highest residuals? Try finding them in the interactive plot by hovering over the dots (HINT: if the residuals are large, they should be far away from the plane, and the sign indicates whether they will be above or below the plane!)

```{r}
sort(resid(reg_full))
```
