---
title: "Lab 7 Activity"
output: pdf_document
header-includes:
    - \usepackage{setspace}
urlcolor: blue
---

\onehalfspacing


We will be looking at the `Chirot` data from the `car` package again. As a reminder, this is data on the [the 1907 Romanian peasant revolt](https://en.wikipedia.org/wiki/1907_Romanian_peasants%27_revolt){target="_blank"}. Find a descriptions of the variables in the table below:




```{r echo=FALSE, message=FALSE, warning=FALSE}

library(car)
library(tidyverse)

info <- tibble(Variable = names(Chirot),
               Description = c("Intensity of the rebellion",
                               "Commercialization of agriculture",
                               "Traditionalism",
                               "Strength of middle peasantry",
                               "Inequality of land tenure"))

library(kableExtra)

kable(info) %>% 
  column_spec(column = 2, width = "300px") %>% 
  row_spec(0, bold = T) %>% 
  column_spec(1, bold = T)
```


run the following code to name the data you will be using as `dat`:

```{r message=FALSE, warning=FALSE}
library(car)
dat <- Chirot
```

**1.** Treat `intensity` as the outcome and all the other variables in the dataset as predictors and run a multiple regression. What is the $R^2$ of this model?

```{r include=FALSE}
reg <- lm(intensity ~ commerce + tradition + midpeasant + inequality, data = Chirot)

summary(reg)$r.squared
```

**2.** We want to rank the predictors in order of importance based on their unique contribution on the total $R^2$. Run a dominance analysis. How much does each predictor contribute to the total $R^2$? (HINT: use the `$` operator on the `dominanceanalysis` object to extract each predictor's contribution without using the `sumamry()` function)

```{r include=FALSE}
library(dominanceanalysis)

DA <- dominanceAnalysis(reg)

DA$contribution.average


```
-  Does `inequality` conditionally dominate `midpeasant`? Motivate your answer based on the appropriate dominance matrix.


```{r include=FALSE}
dominanceMatrix(DA, type = "conditional")
# Yes (1 in the row)
```

-  Does `inequality` completely dominate `midpeasant`? Motivate your answer based on the appropriate dominance matrix.

```{r include=FALSE}
dominanceMatrix(DA, type = "complete")

# inconclusive (.5 in the row)
```
**3.** Although `tradition`, on average, was the second predictor that contributed the most to $R^2$, it was not significant in the full regression. Conduct a retrospective power analysis to evaluate how much power there was to detect a significant effect of `tradition`. Follow the steps below:

- Calculate the $\Delta R^2$ between the full regression and the regression without `tradition`.

```{r include=FALSE}

reg_simp <- lm(intensity ~ commerce + midpeasant + inequality, data = Chirot)

DR2 <- summary(reg)$r.squared - summary(reg_simp)$r.squared

summary(reg_simp)


anova(reg_simp, reg)
```

- calculate $f^2$.

```{r include=FALSE}
f2 <- DR2/(1 - summary(reg)$r.squared)
```

- calculate power (you need to specify the correct $\mathrm{df}_2$ through the `v =` argument and leave the `power =` argument empty).

```{r include=FALSE}
pwr::pwr.f2.test(u = 1,  
                 v = 27,
                 f2 = f2, 
                 sig.level = 0.05)
# power = 0.4854163
```

