---
title: "Lab 3 Activity"
output: pdf_document
header-includes:
    - \usepackage{setspace}
urlcolor: blue
---

\onehalfspacing

```{r include=FALSE}
library(tidyverse)
```

For this activity we will be looking at some data about [the 1907 Romanian peasant revolt](https://en.wikipedia.org/wiki/1907_Romanian_peasants%27_revolt){target="_blank"}! The data is called `Chirot` and is stored in the `carData` package. Run `help("Chirot")` to find out more about the data.


**1.** Load the `carData` and create a separate `data.frame` object called `rev_int` that only contains the the `intensity` and `inequality` variables from the `chirot` data.

```{r include=FALSE}
library(carData)

# MANY ways of doing this

rev_int <- Chirot[,c("intensity","inequality")]

# or also any of tyhe lines below works

# rev_int <- Chirot[,c(1,5)]
# rev_int <- dplyr::select(.data = Chirot, intensity, inequality)
# rev_int <- data.frame(Chirot$intensity, Chirot$inequality)
# rev_int <- cbind(Chirot$intensity, Chirot$inequality)

```

**2.** Find the correlation between `inequality` and `intesity`. Is it significantly different from 0? What is the 95% confidence interval?

```{r include=FALSE}

cor <- psych::corr.test(rev_int)

# correlation
cor$r[2]

# significance and 95%CI
cor$ci

# not significant
```


**3.** Run a linear regression with `inequality` predicting `intensity`. Report the estimated regression coefficient in APA style. Do you notice any similarities between the regression coefficient of `inequality` and the correlation you calculated in the previous question? (yes, what is the same?)

```{r include=FALSE}
reg <- lm(intensity ~ inequality, rev_int)

summary(reg)
```


**4.** According to the previous regression model, what happens to the `intensity` once `inequality ` increases by 3 units?

```{r include=FALSE}
# intensity increases by

coef(reg)[2]*3
```

**5.** What is the predicted value of `intensity` if `inequality` is 10? Make sure there is no rounding error in the answer (i.e., do not copy and paste numbers from the output).

```{r include=FALSE}
coef(reg)[1] + coef(reg)[2]*10  
```
**6.** Create a QQplot that evaluates the normality of residuals. What do you think?

```{r include=FALSE}

# not doing very well

car::qqPlot(reg, id = FALSE)
```

- Afterwards, create a scatterplot with `inequality` on the x-axis and `intensity` on the y-axis to check whether a linear relationship is appropriate for the two variables. What do you think?

```{r include=FALSE}
library(tidyverse)

ggplot(rev_int,
       aes(x = inequality, y = intensity)) +
     geom_point() +
  geom_smooth(method = "loess",
              se = FALSE) +
  theme_classic()

# the relationship between the two variables is not linear. Linear regression does not characterize the relationship between the two variables well.

# Likely poisson/negative binomial process? (https://en.wikipedia.org/wiki/Poisson_distribution) (https://en.wikipedia.org/wiki/Negative_binomial_distribution)
```



\newpage


## Some R Practice: `for` loops

Loops are used to repeat an operation a certain number of times based on certain criteria. There are different types of loops, but the most used type of loop across computer languages is the `for` loop. 

Here is an example of a `for` loop:

```{r}
for(i in 1:10){
  
  print(2*i)
}
```
Did you see what happened? the loop ran *10 times*, determined by the `1:10` part. Remember that `1:10` actually represents:

```{r}
1:10
```

Now, the loop was repeated 10 times, but every time `i`, the *index*, was substituted for one of the numbers in `1:10`. So every time the loop multiplied 2 by a number from 1 to 10. 

This is a simple example, but loops are used in many ways. One way that loops are often used is to save a series of results to some object. The task requires you to do one such operation:\.


**Task**

- You have the following matrix `dat_ex` and the empty object `SDs`. 

```{r}
dat_ex <- cbind(c(1,4,6,77,3,12),
                c(87,789,35,72,33,4),
                c(134,455,66,776,397,1289))

# create an empty object

SDs <- c()
```


Create a `for` loop that stores the standard deviations of each of the columns of `dat_ex` into the `SDs` object. At the end of the `for` loop, the `SDs` object should look like this:

```{r include=FALSE}
# ncol() is "better" than just using 3 because the code will generalize to any number of columns

for(i in 1:ncol(dat_ex)){
  
  SDs[i] <- sd(dat_ex[,i])
  
}
```


```{r}
SDs
```

To solve this task, it is important to know how to index vectors and matrices as discussed in the R practice from Lab 1. 

**Note:** use the `sd()` function calculate the standard deviation. 











