---
title: "Lab 11 Activity"
output: pdf_document
header-includes:
    - \usepackage{setspace}
---


\onehalfspacing

we will continue to look at salaries of university professors in 2008:

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(carData)
library(tidyverse)

dat <- Salaries


info <- tibble(Variable = colnames(dat),
               Description = c("Assistant, Associate, or Full Professor",
                               "A = theoretical department, B = applied department",
                               "Years since PhD",
                               "Years of Service",
                               "Sex (Male or Female)",
                               "nine-moth salary in dollars"))

library(kableExtra)

kable(info) %>% 
  column_spec(column = 2, width = "300px") %>% 
  row_spec(0, bold = T) %>% 
  column_spec(1, bold = T)
```


Run the following code to name the data you will be using as `dat`:

```{r}
library(carData)

dat <- Salaries
```

**1.** We want to test whether there is a difference in salaries (`salary`) between theoretical and applied departments (`discipline`). However, we also believe that this difference in salary should be moderated by how long someone has been at a certain department (`yrs.service`). Run a regression that tests these hypotheses.

```{r include=FALSE}
reg <- lm(salary ~ discipline*yrs.service, dat)
summary(reg)
```

- before interpreting the results, do you believe it meaningful to interpret the regression without centering `yrs.service`? Explain. 

```{r include=FALSE}
# in this case it should be fine to leave the variable as is because 0 years of service is meaningful (i.e., a professor who has just started)
```


- Is the interaction term significant? What can you conclude about the difference in the relation between years of service and salary depending on the type of department?

```{r include=FALSE}
# the interaction term is significant, meaning that the salary of professors in applied departments increases at a faster rate as years of service increase compared to professors in a theoretical department.
```


**2.** Calculate the expected value of `Salary` for a professor in a theoretical department (`discipline` = A) with 15 years of service. 

```{r include=FALSE}
coef(reg)[1] + coef(reg)[3]*15
```
- Now do the same for a professor in an applied field (`discipline` = B).

```{r include=FALSE}
coef(reg)[1] + (coef(reg)[3] + coef(reg)[4])*15 + coef(reg)[2] 
```

**3.** Test whether predicted salaries of professors in applied and theoretical departments differ significantly at different values of years fo service. Test this at `yrs.service` = 0, 15, 30 and 45.

```{r include=FALSE}
mylist <- list(yrs.service=c(0, 15, 30, 45))

means <- emmeans::emmeans(reg, 
                          ~yrs.service*discipline, 
                          at = mylist)

emmeans::contrast(means, "pairwise", by= "yrs.service")
```

- What pattern do you observe as `yrs.service` increases? 

```{r include=FALSE}
# as years of service increases the gap in salary between professors in applied and theoretical departments increases, with professors in applied departments making more and more money
```

- plot the simple slopes for professors in applied and theoretical departments. Do the simple slopes confirm your observation in the previous question?

```{r include=FALSE}
# 0 is the minimum and 60 is the maximum

mylist <- list(yrs.service=c(0, 60))

emmeans::emmip(reg, discipline~yrs.service, at = mylist)

```
```{r include=FALSE}
# Yes, salary of professors in applied departments increases at a much greater rate than the salary of professors in theoretical departments
```




