---
title: "Lab 1 Activity"
output: pdf_document
header-includes:
    - \usepackage{setspace}
---


\onehalfspacing

## Lab Activity 

```{r include=FALSE}
library(tidyverse)
```


**1.** Read "NY_Temp.txt" into R as shown in Lab 1. 

```{r include=FALSE}
NY_temp <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/NY_Temp.txt")
```

**2.** Use a function to find the descriptive statistics including IQR for the windspeed variable, `Wind`.

```{r include=FALSE}
psych::describe(NY_temp$Wind, IQR = TRUE)
```

**3.** Make a basic index plot (scatterplot of Windspeed according to Case Number).

```{r include=FALSE}
ggplot(NY_temp,  
 aes(x = Case, y = Wind)) +
 geom_point() +
 theme_classic()
```

**4.** Add a linear regression line and a quadratic regression line to the index plot. Which line describes the trend more appropriately in your opinion?

```{r include=FALSE}
ggplot(NY_temp,  
 aes(x = Case, y = Wind)) +
 geom_point() +
 theme_classic() +
 geom_smooth(method = "lm", 
             se = FALSE) +
 geom_smooth(method = "lm", 
             formula = y ~ poly(x, 2), 
             color = "red", 
             se = FALSE)

#based on the graph, either solution might be appropriate. I would personally choose linear because it's close enough and makes interpretation easier, but other researchers might choose quadratic and that's ok.
```

**5.** Create a 1D-scatterplot of Wind.

```{r include=FALSE}
ggplot(NY_temp,  
       aes(x = Wind, y = 0)) +
       geom_point(shape = 1) +
       theme_classic()
```

**6.** Choose and create one graph to help you evaluate normality (histogram, boxplot, Q-Q plot, or kernal density plot).

```{r include=FALSE}
ggplot(NY_temp,
       aes(x = Wind)) +
       geom_density() +
       theme_classic()
```




\newpage


## Some R Practice

**1.** The `rep()` function can be used to repeat objects:

```{r}

# we define vectors with the `c()` function

vec <- c(2, 3)
vec

# can you spot the difference between the `times =` and `each =` argument?  

rep(vec, times = 3)

rep(vec, each = 3)
```

**Task**

- Repeat the vector `c(-1,3,-5,7,-9)` twice, with each element repeated 3 times, and store the result. Display the result sorted from largest to smallest (use the `sort()` function; run `help(sort)` for the help page of the `sort()` function).

```{r include=FALSE}
x <- c(-1,3,-5,7,-9)

x_rep <- rep(rep(x, each = 3), 2)

sort(x_rep, decreasing = TRUE)
```

**2.** 

**Task**

-  If I have a vector with 5 elements, I use the `rep()` function to repeat each element 4 times, and then I repeat the resulting vector 3 times, how long is the final vector?

```{r include=FALSE}
5*4*3

# demonstration
vec <- c(1, 1, 1, 1, 1)

length(rep(rep(vec, each = 4), times = 3))

```

**3.**  Let's look at some common R objects and how to subset them:

vectors (1-Dimensional):

```{r}
x <- c(2,1,5,6, 17)

# element 3 is extracted like so

x[3]

# elements 1 and 3 can be extracted like so

x[c(1,3)]

# elements can be dropped by using the "-" sign
# to drop the elements 2 and 4

x[-c(2,4)]
```

matrices (2-Dimensional):

```{r}
# many ways to define matrices
# a quick way is to use the `cbind()` of `rbind()` functions
# `cbind()` glues together 1D vectors as columns 
# `rbind()` glues together 1D vectors as rows 

# for example we create a matrix we 3 rows and 4 columns

mat <- rbind(c(4,2,5, 8),
             c(14,78,6, 38),
             c(33,7,10, 326))

# to extract the element in the second row and the fourth column, we run

mat[2, 4]

# so for 2D objects, [Row, Column]

# to get the second column

mat[,2]

```

lists (1-Dimensional):

```{r}
# lists can store any type of object
# to create a list

list_1 <- list(x, mat)

# to extract elements from a list we use [[]]
# so to extract the second element

list_1[[2]]

# we can also name list elements

list_names <- list("vector" = x,
                   "matrix" = mat)

# now we can also use the `$` operator to access named elements

list_names$matrix

# the [[]] method still works just fine

list_names[[2]]
```

**Tasks**

- If you create a new vector `y` that contains the first to the fourth element of `x` and the third row of the `mat` object, what will be the length of `y`?

```{r include=FALSE}
length(x[1:4]) + length(mat[3,]) 
```


- Create `y` as specified above. Print `y`.

```{r include=FALSE}
y <- c(x[1:4], mat[3,])
y
```

- Extract the second and fifth elements of the "vector" element inside the `list_names` object in a single line of code

```{r include=FALSE}
list_names$vector[c(2,5)]
```
- assign `NA` to the element in the third row and second column of the "matrix" element inside the `list_names` object in a single line of code. (hint: you will need to use the `<-` operator). Print the "matrix" element inside the `list_names` object to confirm.

```{r include=FALSE}
list_names$matrix[3, 2] <- NA

list_names$matrix
```














