---
title: "Lab 10: Categorical Predictors"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
notice: |
  @Wickham_RStudio_2023
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"

format:
   revealjs:
      footer: "Lab 10: Categorical Predictors"
      width: 1280
      height: 720
      chalkboard: true
      revealjs-fontawesome: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

  
editor: source
---

## Font-awesome slide{visibility="hidden"}

uncounted and not visible slide to trigger the fontawesome extension

{{< fa thumbs-up >}} 


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

# install.packages("tidyverse")
install.packages("lubridate")
install.packages("splines2")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(tidyverse)
theme_set(theme_classic(base_size = 14, 
                        base_family = 'serif'))
library(lubridate)
library(splines2)
```


</br>

<div style="font-size: 26px">

::: {.panel-tabset}
### `lubridate`

The `lubridate` package [@Spinu_etal_2024] provides many functions to help you work with date and times.

### `splines2`

The `splines2` package [@Wang_etal_2024] can be used to calculate splines. 

:::

</div>


:::


::: {.column width="50%"}

<center style="padding-bottom: 41px;"> [Data]{.data-title} </center>

<div style="font-size: 24px"> This data come from a random sample of 200 students from the 1982 [high-school and beyond longitudinal study.](https://nces.ed.gov/surveys/hsb/){target="_blank"} </div>



```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

hsb2 <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/hsb2.csv")

```
```{r}
reactable::reactable(hsb2,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1.175rem"),
                     pagination = FALSE, highlight = TRUE, height = 350)
```



:::
::::





## References 

<div id="refs"> </div>
