---
title: "Lab 7: Multicollinearity, Suppression, Dominance Analysis"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
notice: |
  @Fox_etal_2024
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "Lab 7: Multicollinearity, Suppression, Dominance Analysis"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---



## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

install.packages("dominanceanalysis")
# install.packages("tidyverse")

```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(dominanceanalysis)
library(tidyverse)
theme_set(theme_classic(base_size = 14, 
                        base_family = 'serif'))
```

</br>

<div style="font-size: 26px">

::: {.panel-tabset}

### `dominanceanalysis`

The `dominanceanalysis` package [@Navarrete_etal_2024] contains functions to run dominance analysis in R.
:::

</div>

:::
::: {.column width="50%"}

<ul>

<li> Let's also load the [data](https://library.virginia.edu/data/articles/hierarchical-linear-regression){target="_blank"}
 for today: </li>
</ul>

```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

# Data from https://library.virginia.edu/data/articles/hierarchical-linear-regression

dat <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/Hierarc_data.csv")


# let's peak at our variables
str(dat, vec.len = 2)
```

:::
::::


## Regresssion coefficients Weirdness

It is common practice to add covariates to regression models without giving it much thought (*adding a covariate* simply means adding additional predictors). However, some strange things 👽 may happen if you add variables too casually to regression models.

:::: {.columns}
::: {.column width="50%"}



:::
::: {.column width="50%"}

:::
::::



## Visulazing 

```{r}
source("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/R_scripts/3D_plot.R")

# dat <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/training.sav")
# 
# 
# cor(dat)
```





## References 

<div id="refs"> </div>


