---
title: "Lab 11: Interactions With Categorical Predictors"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "Lab 11: Interactions With Categorical Predictors"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

# install.packages("tidyverse")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(tidyverse)
theme_set(theme_classic(base_size = 14, 
                        base_family = 'serif'))
library(fastDummies)
```


</br>

<div style="font-size: 26px">

::: {.panel-tabset}
### `fastDummies`

<div style="font-size: 22px"> The `fastDummies` package [@Kaplan_Schlegel_2025] provides some functions to help with dummy coding. Later, I show a better way of achieving dummy coding in `R`, but this package offers the "more standard" procedure. </div>

:::

</div>

  

:::


::: {.column width="50%"}

<center style="padding-bottom: 41px;"> [Data]{.data-title} </center>

<div style="font-size: 22px"> We'll continue with the `hsb2` dataset from [Lab 10](https://quinix45.github.io/PSYC-7804-Regression-Lab-Slides/Slides%20Files/Lab%2010.html#/todays-packages-and-data){target="_blank"}. 
 </div>
 
 
```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

hsb2 <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/hsb2.csv")
```

<center style="padding-bottom: 11px;"> </center>

```{r}
reactable::reactable(hsb2,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1.105rem"),
                     pagination = FALSE, highlight = TRUE, height = 300)
```

:::
::::



## What are we looking at Today?

In Lab 10 we looked at categorical predictors and a bunch of coding schemes for categorical variable, but we will stick to [dummy coding](https://quinix45.github.io/PSYC-7804-Regression-Lab-Slides/Slides%20Files/Lab%2010.html#/creating-dummy-coded-columns){target="_blank"} for this lab.

We eventually want to look at whether `gender` ($Z$) moderates how `socst` ($X$) score predicts `write` ($Y$) score. As a reminder:


:::: {.columns}
::: {.column width="50%"}

<ul style="font-size: 24px">  

<li> `gender`: student gender (either *male* or *female*).  </li>
<li> `socst`: score on a standardized social studies assessment. </li>
<li> `write`: score on a standardized writing assessment. </li>

</ul>

:::


::: {.column width="50%"}
`gender` is a categorical variable, so let's turn it into a `factor`

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

hsb2$gender <- factor(hsb2$gender)

contrasts(hsb2$gender)
```


:::
::::

So for the dummy coded `gender` variable, $0 =$ *female* and $1 = $ *male* 
  




```{r}

# reg_int <- lm(write ~ scale(socst, center = TRUE)*gender ,data = hsb2)
# reg_int <- lm(write ~ scale(socst, center = TRUE)*schtyp ,data = hsb2)
# 
# summary(reg_int)
```




## References 

<div id="refs"> </div>







