---
title: "Lab 8: Quadratic regression and non-linear alternatives"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
notice: |
  @Fox_etal_2024
  @Wickham_RStudio_2023
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "Lab 8: Quadratic regression and non-linear alternatives"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---


## Today's Packages and Data ðŸ¤—

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

# install.packages("tidyverse")
install.packages("lubridate")
install.packages("splines2")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(tidyverse)
theme_set(theme_classic(base_size = 14, 
                        base_family = 'serif'))
library(lubridate)
library(splines2)
```


</br>

<div style="font-size: 26px">

::: {.panel-tabset}
### `lubridate`

The `lubridate` package [@Spinu_etal_2024] provides many functions to help you work with date and times.

### `splines2`

The `splines2` package [@Wang_etal_2024] can be used to calculate splines. 

:::

</div>


:::


::: {.column width="50%"}

<center style="padding-bottom: 41px;"> [Data]{.data-title} </center>


```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

NY_waeth <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/NYC_Weather_2016_2022.csv")

str(NY_waeth, vec.len = 2)
```

:::
::::


::: {.fragment fragment-index=1}
The data includes weather recordings in NY for every hour of the day over a period of roughly 7 years (lots of data ðŸ˜±). This is a bit too granular for what we need to do, so let's do some data cleaning!
:::


## Working with Dates

This Data provides a good opportunity to show how you can work with dates in R! Although I will only show one function here, the `lubridate` package is your friend if you work a lot with dates and times in your research. 

:::: {.columns}
::: {.column width="50%"}


::: {.fragment fragment-index=1}
<div style="font-size: 22px"> We want to turn the `time` variable into a date in days that R recognizes. we would use the `ymd()` function normally, but the correct format should be *YYYY-MM-DD*. Instead we also have the time here, which we do not need: </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

NY_waeth$time[1:3]
```
:::

::: {.fragment fragment-index=2}

<div style="font-size: 22px; padding-top: 12px;"> We can do some [string manipulation](https://www.geeksforgeeks.org/string-manipulation-in-r/){target="_blank"} to remove the last 6 characters! </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# substr() takes a sting and extract characters from (1 here) to (10 here)

NY_waeth$time <- substr(NY_waeth$time, 1, 10)
NY_waeth$time[1:3]
```
:::


:::
::: {.column width="50%"}


::: {.fragment fragment-index=3}

<div style="font-size: 22px"> Now we can turn the `time` variable into a date: </div>


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

NY_waeth$time <- lubridate::ymd(NY_waeth$time)
class(NY_waeth$time)
```
:::


::: {.fragment fragment-index=4}
<div style="font-size: 22px; padding-top: 12px;"> It is interesting to note that dates in R are interpreted as the time that has passed since the [Unix Epoch](https://en.wikipedia.org/wiki/Unix_time){target="_blank"}, January 1<sup>st</sup>, 1970. This is how most software/computers keep track of time ðŸ¤“
  </div>


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

NY_waeth$time[1]
as.numeric(NY_waeth$time[1])/365
```

<div style="font-size: 24px; padding-top: 12px;"> So "2016-01-01" is 46 years after January 1<sup>st</sup>, 1970. </div>

:::


:::
::::

## Summarizing and Filtering the Data

<div style="font-size: 22px; padding-bottom: 14px;"> We still have 23 measurements per day over 7 years, which is a bit much. For now, let's reduce our data to just the average temperature each day in 2021. </div>

:::: {.columns}
::: {.column width="50%"}

::: {.fragment fragment-index=1}

<div style="font-size: 22px"> We will be using some functions from the `dplyr` package, which is always loaded by `tidyverse`. Let's get the average for each day: </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


Temp_avg <- NY_waeth %>% 
              select(time, `temperature_2m (Â°C)`) %>%
                group_by(time) %>% 
          summarise(Temp_C = mean(`temperature_2m (Â°C)`))

str(Temp_avg)
```
:::


::: {.fragment fragment-index=2}

::: {.callout-note}

## The `tidyverse` Pipe Operator %>% 

The `%>%` operator, AKA the *pipe operator*, passes the element on its left to the function to its right. It is helpful to do multiple operations without needing to redefine an object on every line. 

::: 
:::
:::


::: {.column width="50%"}

::: {.fragment fragment-index=3}

<div style="font-size: 22px"> For the Americans ðŸ¦…, let's also transform temperature to Fahrenheit (The inferior way of measuring temperature ðŸ˜¤) </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


Temp_avg$Temp_F <- (Temp_avg$Temp_C * 9/5) + 32 
```

:::


::: {.fragment fragment-index=4}
<div style="font-size: 22px; padding-top: 14px;"> Now let's filter the data fo the year 2021. We can take advantage of the fact that Dates in R are treated as numbers: </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


Temp_avg_2021 <- Temp_avg %>% 
                  filter(time >= "2021-01-01", 
                         time <= "2021-12-30")
str(Temp_avg_2021)
```

</br>

<center> âœ¨ clean data âœ¨</center>
:::


:::
::::

## Plot The Data

Let's see what is the temperature trend in NY for the year 2021.

:::: {.columns}
::: {.column width="30%"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
#| classes: code-125

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point()
```

Quite clearly, we need a quadratic curve here.

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point()

```

:::
::::


## Plot The Data 

Let's see what is the temperature trend in NY for the year 2021.

:::: {.columns}
::: {.column width="30%"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
#| classes: code-125

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point() +
geom_smooth(method = "lm", 
            formula = y ~ poly(x, 2), 
            se = FALSE)
```

The quadratic line fits very well. Let's see if the loess line agrees.

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point()  +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE)
  

```

:::
::::

## Plot The Data 

Let's see what is the temperature trend in NY for the year 2021.

:::: {.columns}
::: {.column width="30%"}

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
#| classes: code-125

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point()  +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE) +
geom_smooth(method = "loess", 
            col = "red",
           se = FALSE)
```

The loess line is pretty close to the quadratic one! 


Let's run a quadratic regression with `time` as the predictor...

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point()  +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE) +
geom_smooth(method = "loess", 
            col = "red",
           se = FALSE)
  
```

:::
::::

## Quadratic Regression: Centering Variables

<div style="font-size: 24px; padding-bottom: 12px;"> I will first show the "manual" way of running a quadratic regression, and I will show the slightly more practical way later (if you are not interested in interpreting coefficients). </div>

:::: {.columns}
::: {.column width="50%"}


::: {.fragment fragment-index=1}

<div style="font-size: 24px"> First, let's create the quadratic version of the `time` variable </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# dates don't like the "^", so we need to turn it into a numeric first 

Temp_avg_2021$time2 <- as.numeric(Temp_avg_2021$time)^2

```
:::


</br>

::: {.fragment fragment-index=2}
<div style="font-size: 24px"> Oh wait...this does not work too well ðŸ«¤ </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


cor(Temp_avg_2021$time2, as.numeric(Temp_avg_2021$time))
```


 We will not be able to interpret our regression coefficients (the linear term specifically) with such multicollinearity, as discussed in [Lab 7](https://raw.githack.com/quinix45/PSYC-7804-Regression-Lab-Slides/main/Slides%20Files/Lab%207.html#/another-perspective-residuals){target="_blank"}
:::

::: {.fragment fragment-index=3}
<center> ðŸ’¡ We need to center our variables first ðŸ’¡ </center>
:::

:::
::: {.column width="50%"}


::: {.fragment fragment-index=4}
To center a variable we just subtract the mean

$$X_{\mathrm{centered}} = X - \bar{X}$$
:::


::: {.fragment fragment-index=5}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

Temp_avg_2021$time_cent <- Temp_avg_2021$time - mean(Temp_avg_2021$time)
# btw, check out what the time_cent variable looks like now!

# we create the quadratic term by squaring the centered variable
Temp_avg_2021$time_cent2 <- as.numeric(Temp_avg_2021$time_cent)^2
```

And the mulitcollinearity problem is solved ðŸª„

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

cor(Temp_avg_2021$time_cent2, 
    as.numeric(Temp_avg_2021$time_cent))
```

<center> Now we are good to proceed with our quadratic regression. </center>

:::

:::
::::


## Quadratic Regression


:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


quad_reg <- lm(Temp_F ~ time_cent + time_cent2 , 
               data = Temp_avg_2021)

summary(quad_reg)
```



:::
::: {.column width="50%"}

::: {.fragment fragment-index=1}

$\widehat{\mathrm{F_{temp}}} = b_0 + b_1 \times \mathrm{time} + b_2 \times \mathrm{time}^2$

<center> &darr; </center>

$\widehat{\mathrm{F_{temp}}} = 71.16 + 0.06 \times \mathrm{time} - 0.001 \times \mathrm{time}^2$

:::

<ul style="font-size: 22px">  


::: {.fragment fragment-index=1}
<li> 71.16 is the expected temperature in Fahrenheit when `time` is 0. Our variable is centered, so 0 represent the middle of the year, roughly July 2<sup>nd</sup> </li>
:::

::: {.fragment fragment-index=2}
<li>  0.06 is the expected slope of the tangent line when `time` is 0 (confusing ðŸ˜•)  </li>
:::

::: {.fragment fragment-index=3}
<li>  -0.001 is the expected change in the slope of the tangent line per unit-change in `time` (even more confusing ðŸ˜•ðŸ˜•) </li>
:::


::: {.fragment fragment-index=4}
We need some graphical help because some of this stuff honestly makes no sense in words ðŸ¤·
:::


</ul>



:::
::::


## $b_0$ in Quadratic Regression


:::: {.columns}
::: {.column width="30%"}

```{r}
#| eval: false
#| echo: true 
#| code-fold: true
#| code-summary: "Plot code"
#| code-line-numbers: false


ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
  geom_point( aes(y = coef(quad_reg)[1], 
                  x = mean(Temp_avg_2021$time)),
              size = 3.5,
              col = "red") +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE) +
geom_segment(x = 0, 
               xend = mean(Temp_avg_2021$time), 
               y = coef(quad_reg)[1],
               yend = coef(quad_reg)[1],
               linetype = 2) +
geom_segment(x = mean(Temp_avg_2021$time), 
             xend = mean(Temp_avg_2021$time), 
             y = 0,
             yend = coef(quad_reg)[1],
             linetype = 2)
```


$$b_0 = 71.16$$

<div style="font-size: 24px"> Given that the `time` variable was centered, 0 corresponds to the middle of the year, July 2<sup>nd</sup>. 



The expected temperature on July 2<sup>nd</sup> (when `time` = 0) given the quadratic regression equation is indeed 71.16 degrees Fahrenheit. </div>

:::
::: {.column width="70%"}


```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
  geom_point( aes(y = coef(quad_reg)[1], 
                  x = mean(Temp_avg_2021$time)),
              size = 3.5,
              col = "red") +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE) +
geom_segment(x = 0, 
               xend = mean(Temp_avg_2021$time), 
               y = coef(quad_reg)[1],
               yend = coef(quad_reg)[1],
               linetype = 2) +
geom_segment(x = mean(Temp_avg_2021$time), 
             xend = mean(Temp_avg_2021$time), 
             y = 0,
             yend = coef(quad_reg)[1],
             linetype = 2)
```


:::
::::

## $b_1$ in Quadratic Regression

:::: {.columns}
::: {.column width="30%"}

```{r}
#| eval: false
#| echo: true 
#| code-fold: true
#| code-summary: "Plot code"
#| code-line-numbers: false


ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
geom_smooth(method = "lm", 
            formula = y ~ poly(x, 2), 
            se = FALSE) +
geom_segment(x = mean(Temp_avg_2021$time) - 130, 
             xend = mean(Temp_avg_2021$time) + 130, 
             y = coef(quad_reg)[1] - coef(quad_reg)[2]*130,
             yend = coef(quad_reg)[1] + coef(quad_reg)[2]*130,
             linetype = 1,
             linewidth = .7,
             col = "red")

```


$$b_1 = 0.06$$

<div style="font-size: 24px"> The slope of the tangent line when `time` is at 0. The tangent line is the red line in the plot on the right. 



For this to make a bit more sense, we also need to look at $b_2$. </div>

::: {.fragment fragment-index=1}

::: {.callout-note}

## Slope-ception ðŸ¤”

If you think about it, in linear regression, the slope, $b_1$, is also the slope of the tangent line to the regression line. 

::: 
:::

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
geom_smooth(method = "lm", 
            formula = y ~ poly(x, 2), 
            se = FALSE) +
geom_segment(x = mean(Temp_avg_2021$time) - 130, 
             xend = mean(Temp_avg_2021$time) + 130, 
             y = coef(quad_reg)[1] - coef(quad_reg)[2]*130,
             yend = coef(quad_reg)[1] + coef(quad_reg)[2]*130,
             linetype = 1,
             linewidth = .7,
             col = "red")

```

:::
::::

## $b_2$ in Quadratic Regression

:::: {.columns}
::: {.column width="30%"}


```{r}
#| eval: false
#| echo: true 
#| code-fold: true
#| code-summary: "Plot code"
#| code-line-numbers: false

x_val <- 100

# accurate calculations, there is rounding error in the equation on the slides
slope_new <- coef(quad_reg)[2] + 2*coef(quad_reg)[3]*x_val

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
geom_smooth(method = "lm", 
            formula = y ~ poly(x, 2), 
            se = FALSE) +
geom_segment(x = (mean(Temp_avg_2021$time) + x_val) - 130, 
             xend = (mean(Temp_avg_2021$time) + x_val) + 130, 
             y =  (coef(quad_reg)[1] + coef(quad_reg)[2]*x_val + coef(quad_reg)[3]*x_val^2) - (slope_new*130),
             yend = (coef(quad_reg)[1] + coef(quad_reg)[2]*x_val + coef(quad_reg)[3]*x_val^2) + (slope_new*130),
             linetype = 1,
             linewidth = .7,
             col = "red") 

```


$$b_2 = - 0.001$$

<div style="font-size: 24px"> $b_2$ lets us calculate the slope of the tangent line at *any* point of the curve. The value of $b_2$, describes the change of the slope of the tangent line per 1-unit change in `time`. More formally, 


$$b_{\mathrm{tang}} = b_1 + 2b_2 X$$ </div>

::: {.fragment fragment-index=1}
<div style="font-size: 22px"> For example, if `time` = 100, then the slope of the tangent line is given by $b_{\mathrm{tang}} = .06 + 2\times -.001 \times100 = -.14$. Indeed, the tangent line goes downwards, as its slope suggests. </div>
:::

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false



x_val <- 100

# accurate calculations, there is rounding error in the equation on the slides
slope_new <- coef(quad_reg)[2] + 2*coef(quad_reg)[3]*x_val

ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
geom_smooth(method = "lm", 
            formula = y ~ poly(x, 2), 
            se = FALSE) +
geom_segment(x = (mean(Temp_avg_2021$time) + x_val) - 130, 
             xend = (mean(Temp_avg_2021$time) + x_val) + 130, 
             y =  (coef(quad_reg)[1] + coef(quad_reg)[2]*x_val + coef(quad_reg)[3]*x_val^2) - (slope_new*130),
             yend = (coef(quad_reg)[1] + coef(quad_reg)[2]*x_val + coef(quad_reg)[3]*x_val^2) + (slope_new*130),
             linetype = 1,
             linewidth = .7,
             col = "red") 
```

:::
::::

## Why $b_2$ is What You care about

So why do we care about slopes of tangent lines? Well, generally, we don't. The thing that we care about is that the slope of tangent lines *changes* at all. This is what a non-zero $b_2$ value tells us.  

::: {.fragment fragment-index=1}
What would it mean for the slope of the tangent line to be *constant across every $X$ value*? What happens if $b_2 = 0$?
:::

::: {.fragment fragment-index=2}
We can check our quadratic regression equation, so $\hat{Y} = b_0 + b_1X +  0X^2$ &rarr; $\hat{Y} = b_0 + b_1X$
:::


::: {.fragment fragment-index=3}
Oh, we are back to the linear regression formula ðŸ™ƒ If the slope of the tangent line is always constant, we have a *line*.
:::

::: {.fragment fragment-index=4}
So, $b_2$ informs us about 3 things:
:::

:::: {.columns}

::: {.column width="33%"}

::: {.fragment fragment-index=5}
**1.** Whether the line bends at all (i.e., is $b_2 \neq 0$) 
:::

:::

::: {.column width="33%"}

::: {.fragment fragment-index=6}
**2.** How much the line bends, which it depends on the *magnitude* of $b_2$ (more extreme $b_2$ values imply steeper curves)
:::

:::


::: {.column width="33%"}

::: {.fragment fragment-index=7}
**3.** The direction of the bend. If the sign of $b_2$ is positive, the curve will be a bell, if the sign is negative, it will be a U, as in our case.
:::

:::
::::

## Climbing to the top â˜ï¸ (or bottom ðŸ‘‡)

:::: {.columns}
::: {.column width="40%"}


```{r}
#| eval: false
#| echo: true 
#| code-fold: true
#| code-summary: "Plot code"
#| code-line-numbers: false

# accurate calculations, there is rounding error in the equation on the slides
top <- -1*(coef(quad_reg)[2]/(2*coef(quad_reg)[3]))

Y_val <- (coef(quad_reg)[1] + coef(quad_reg)[2]*top + coef(quad_reg)[3]*top^2)

# this is the formula for the tangent slope. if you print this, you will see that it equals exactly 0. This means that the X value provided is the top of the curve, which is the only point where the tangent slope is exactly 0

slope_new <- coef(quad_reg)[2] + 2*coef(quad_reg)[3]*top


ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
  geom_point(aes(y = Y_val, 
                x = mean(Temp_avg_2021$time) + top),
              size = 3.5,
              col = "red") +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE) +
geom_segment(x = 0, 
               xend = mean(Temp_avg_2021$time) + top, 
               y = Y_val,
               yend = Y_val,
               linetype = 2) +
geom_segment(x = mean(Temp_avg_2021$time) + top, 
             xend = mean(Temp_avg_2021$time) + top, 
             y = 0,
             yend = Y_val,
             linetype = 2) +
geom_segment(x = mean(Temp_avg_2021$time) -130, 
             xend = mean(Temp_avg_2021$time) + 130, 
             y =  Y_val,
             yend = Y_val,
             linetype = 1,
             linewidth = .7,
             col = "red") 
```


<div style="font-size: 24px"> Last but not least, we can find the top (or bottom if $b_2 > 0$) of our curve (which is also the turning point). The formula is $\mathrm{top} = -\frac{b_1}{2b_2}$. </div>


::: {.fragment fragment-index=1}
<div style="font-size: 24px"> For us that would mean that the predicted highest average temperature is around:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# accurate calculations, there is rounding error in slide regression coefficients
-1*(coef(quad_reg)[2]/(2*coef(quad_reg)[3]))
```

July 22<sup>nd</sup> </div>
:::


:::
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false


# accurate calculations, there is rounding error in the equation on the slides
top <- -1*(coef(quad_reg)[2]/(2*coef(quad_reg)[3]))

Y_val <- (coef(quad_reg)[1] + coef(quad_reg)[2]*top + coef(quad_reg)[3]*top^2)

# this is the formula for the tangent slope. if you print this, you will see that it equals exactly 0. This means that the X value provided is the top of the curve, which is the only point where the tangent slope is exactly 0

slope_new <- coef(quad_reg)[2] + 2*coef(quad_reg)[3]*top


ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point(alpha = .5, shape = 1)  +
  geom_point(aes(y = Y_val, 
                x = mean(Temp_avg_2021$time) + top),
              size = 3.5,
              col = "red") +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE) +
geom_segment(x = 0, 
               xend = mean(Temp_avg_2021$time) + top, 
               y = Y_val,
               yend = Y_val,
               linetype = 2) +
geom_segment(x = mean(Temp_avg_2021$time) + top, 
             xend = mean(Temp_avg_2021$time) + top, 
             y = 0,
             yend = Y_val,
             linetype = 2) +
geom_segment(x = mean(Temp_avg_2021$time) -130, 
             xend = mean(Temp_avg_2021$time) + 130, 
             y =  Y_val,
             yend = Y_val,
             linetype = 1,
             linewidth = .7,
             col = "red") 
```

:::
::::



::: {.fragment fragment-index=2}
<div style="font-size: 24px"> Why does $-\frac{b_1}{2b_2}$ always finds the top/bottom? Turns out that $-\frac{b_1}{2b_2}$ finds the $X$ point where the slopes of the tangent line is *exactly 0*. That means that the tangent line is *perfectly flat*, which happens only at the top/bottom of a quadratic function. </div>
:::

::: {.fragment fragment-index=3}
<div style="font-size: 22px; padding-top: 12px;"> This concept of ["finding the top (or bottom)"](https://en.wikipedia.org/wiki/Mathematical_optimization){target="_blank"} happens to be the foundation of almost all estimation algorithms in statistics. </div>
:::

## Bootsrap CI for Maximum/Minimum point

We can create a 95% CI for the `time` value at which we expect the maximum temperature with bootstrapping.

:::: {.columns}
::: {.column width="50%"}


::: {.fragment fragment-index=1}

<div style="font-size: 24px"> First we create the bootstrapped regression: </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

boot_quad <- car::Boot(quad_reg, R = 1000)
```
<div style="font-size: 24px; padding-top: 12px;"> Then we need to calculate $-\frac{b_1}{2b_2}$ for all our 1000 bootstrap samples (very easy actually!). The bootstrap samples are saved in the `t` element of `boot_quad`:  </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# View(boot_quad$t)
# we can compute -b1/(2*b2) for all sample at once with this

top_boot <- -boot_quad$t[,2]/(2*boot_quad$t[,3])
```

:::

:::
::: {.column width="50%"}


::: {.fragment fragment-index=2}

<div style="font-size: 24px; padding-top: 12px;"> The 95% CI for the `time` value at which we expect the highest temperatures is: </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

quantile(top_boot, c(.025, .975))
```
<div style="font-size: 24px;"> And the full distribution </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| out-width: "80%"

plot(density(top_boot))
```
:::

:::
::::


## How Many Polynomials do I need?

It is my opinion that will almost never need more than a quadratic term if you are doing research in Psychology. In our case, we are dealing with a physical phenomenon (temperature), so higher polynomials may not be as far fetched. 


::: {.fragment fragment-index=1}
<div style="font-size: 24px">  How do we find out how many polynomials we need? Since we are just adding additional variables to the regression model, we can use in  $\Delta R^2$ just as we did in [Lab 6](https://raw.githack.com/quinix45/PSYC-7804-Regression-Lab-Slides/main/Slides%20Files/Lab%206.html#/testing-gain-in-prediction-delta-r2){target="_blank"}. </div>
:::


:::: {.columns}
::: {.column width="60%"}


</br>


::: {.fragment fragment-index=2}

<div style="font-size: 24px"> We can easily add polynomials to regressions thanks to the `poly()` function: </div>


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_lin <- lm(Temp_F ~ time, data = Temp_avg_2021)  
reg_quad <- lm(Temp_F ~ poly(time, 2), data = Temp_avg_2021)  
reg_cube <- lm(Temp_F ~ poly(time, 3), data = Temp_avg_2021)  
reg_fourth <- lm(Temp_F ~ poly(time, 4), data = Temp_avg_2021) 
reg_fifth <- lm(Temp_F ~ poly(time, 5), data = Temp_avg_2021) 
```
:::

:::
::: {.column width="40%"}

</br>

::: {.fragment fragment-index=2}

::: {.callout-warning}

## The `poly()` Function 

Be careful! It may seem that the `poly()` function saves us a lot of time, but it calculates **orthogonal polynomials**. Do not interpret the regression coefficients of regressions that use `poly()` functions! For more, see [this link](https://online.stat.psu.edu/stat502/lesson/10/10.2){target="_blank"}.
::: 
:::

:::
::::

## Comparing Polynomial Regressions

:::: {.columns}
::: {.column width="50%"}

The $\Delta R^2$ suggests that there is a significant $R^2$ improvement up to a polynomial of the 4<sup>th</sup> degree. Here are the $R^2$ values of the 5 regression models:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125



c(summary(reg_lin)$r.squared,
  summary(reg_quad)$r.squared,
  summary(reg_cube)$r.squared,
  summary(reg_fourth)$r.squared,
  summary(reg_fifth)$r.squared)

```
The 5<sup>th</sup> degree polynomial offers essentially no improvement. 

:::
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

anova(reg_lin, reg_quad, reg_cube, reg_fourth, reg_fifth)
```

:::
::::

</br>

<center>

::: {.fragment fragment-index=1}
Let's quickly look at what the polynomial regression lines look like.
:::

</center>


## Plotting polynomials


:::: {.columns}
::: {.column width="50%"}

```{r out.width="80%"}
ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point() +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 2), 
           se = FALSE) +
  ggtitle("Second Degree Polynomial") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
  
```


:::
::: {.column width="50%"}

```{r out.width="80%"}
ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point() +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 3), 
           se = FALSE) +
  ggtitle("Third Degree Polynomial") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

:::
::::


:::: {.columns}
::: {.column width="50%"}

```{r out.width="80%"}
ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point() +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 4), 
           se = FALSE) +
  ggtitle("Fourth Degree Polynomial") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

:::
::: {.column width="50%"}

```{r out.width="80%"}
ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point() +
geom_smooth(method = "lm", 
           formula = y ~ poly(x, 5), 
           se = FALSE) +
  ggtitle("Fifth Degree Polynomial") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

:::
::::

## Piecewise Regression

<div style="font-size: 24px">  In piecewise regression we effectively divide our data into sections that we run separate regression on. You can divide your data in however many sections you want, but usually you should not need more than 2 separate segments. </div>


:::: {.columns}
::: {.column width="30%"}

::: {.fragment fragment-index=1}
<div style="font-size: 24px; padding-top: 16px;">  But how do we divide the data? *You* have to decide it based on what *you* know about your data/theory ðŸ˜€ </div>
:::

::: {.fragment fragment-index=2}
<div style="font-size: 22px; padding-top: 16px;"> In the case of this data, I think that a good cut-point is the top of the quadratic regression from before! </div>


<ul style="font-size: 22px">  

<li>Why? Because we know that the top is the turning point of the quadratic line. If I need to fit two lines to the data, that is probably my best option given what I know. </li>

</ul>
:::

::: {.fragment fragment-index=3}
<div style="font-size: 20px;">  The plot on the right shows the segments if we divide the data in this way. </div>
:::


:::
::: {.column width="70%"}

::: {.fragment fragment-index=3}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"

# split data with cut-point (cleaner code in 2 slides, can't be bothered to change it)

Temp_avg_2021$D1 <- ifelse(Temp_avg_2021$time_cent >= 20.34353, 0*Temp_avg_2021$time_cent, 1*Temp_avg_2021$time_cent)
Temp_avg_2021$D2 <- ifelse(Temp_avg_2021$time_cent < 20.34353, 0*Temp_avg_2021$time_cent, 1*Temp_avg_2021$time_cent)



reg_piece <- lm(Temp_F ~ D1 + D2, data = Temp_avg_2021)

intercept <- coef(reg_piece)[1]
D1 <- coef(reg_piece)[2]
D2 <- coef(reg_piece)[3]


ggplot(Temp_avg_2021,
       aes(x = time, y = Temp_F)) +
        geom_point() +
geom_segment(x = min(Temp_avg_2021$time_cent) + mean(Temp_avg_2021$time), 
             xend = 20.34353 + mean(Temp_avg_2021$time), 
             y = intercept + min(Temp_avg_2021$time_cent)*D1 ,
             yend = intercept,
             linetype = 1,
             linewidth = .7,
             col = "red") +
geom_segment(xend = max(Temp_avg_2021$time_cent) + mean(Temp_avg_2021$time), 
             x = 20.34353 + mean(Temp_avg_2021$time), 
             yend = intercept + max(Temp_avg_2021$time_cent)*D2 ,
             y = intercept,
             linetype = 1,
             linewidth = .7,
             col = "red") +  
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

```
:::
:::
::::



## Setting up Piecewise 

There are multiple ways to set up piecewise regression in `R`. I find the following method the most intuitive:


:::: {.columns}
::: {.column width="50%"}

$$
\hat{Y} = b_0 + b_1\times D_1 + b_2 \times D_2
$$


:::
::: {.column width="50%"}

<div style="font-size: 24px"> Where $b_1$ represent the slope of the *first segment* and $b_2$ represents the slope of the *second segment*. </div>

:::
::::

::: {.fragment fragment-index=1}
We need to create $D_1$ and $D_2$. The trick is to get the original predictor, `time_cent` in our case, and *split* its value across the segments.
:::

:::: {.columns}
::: {.column width="50%"}

::: {.fragment fragment-index=2}
So, if we choose our cut-point to be around the top or the quadratic curve, roughly $20.34$:


<ul style="font-size: 22px">  

<li> $D_1:$ Make all values *after* $20.34$ become $0$. </li>
<li> $D_2:$ Make all values *before* $20.34$ become $0$. </li>

</ul>
:::

::: {.fragment fragment-index=3}
We will use the `ifelse()` function to make our life easier. 
:::

:::
::: {.column width="50%"}

::: {.fragment fragment-index=3}
`ifelse()` syntax: `ifelse(CONDITION, if CONDITION TRUE do something, if CONDITION FALSE do something)`

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# an example
ifelse(c(2,3) > 2.5, "bigger", "smaller")
```
:::

:::
::::

## Creating $D_1$ and $D_2$

:::: {.columns}
::: {.column width="55%"}

- Creating $D_1$:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

Temp_avg_2021$D1 <- ifelse(Temp_avg_2021$time_cent >= 20.34353, 
                           0, Temp_avg_2021$time_cent)
```

- Creating $D_2$:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

Temp_avg_2021$D2 <- ifelse(Temp_avg_2021$time_cent < 20.34353, 
                           0, Temp_avg_2021$time_cent)
```


::: {.fragment fragment-index=1}
<div style="font-size: 24px; padding-top: 16px;"> If you look at our variables on the left and scroll until July 22<sup>nd</sup>, you should see how the `temp_cent` variable has been split between $D_1$ and $D_2$.
 </div>
:::


:::
::: {.column width="45%"}

::: {.fragment fragment-index=1}
```{r}
reactable::reactable(Temp_avg_2021[, c("time", "time_cent", "D1", "D2")],
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "0.875rem"),
                     pagination = FALSE, highlight = TRUE, height = 500)
```
:::

:::
::::

## Running The Piecewise Regression

Now we can simply run our piecewise regression by using $D_1$ and $D_2$ as predictors:

:::: {.columns}
::: {.column width="70%"}


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

piece_reg <- lm(Temp_F ~ D1 + D2, data = Temp_avg_2021)
summary(piece_reg)
```
:::
::: {.column width="30%"}

<div style="font-size: 24px">  
 As we saw from the plot some slides ago, the slope of the first segment, $D_1$ is positive, and that slope of the second segment, $D_2$ is negative. $R^2$ is also quite good at $81\%$.</div>


 <div style="font-size: 24px; padding-top: 12px;"> You can interpret these slopes as you would normal regression slopes. Just remember where $D_1$ ends and $D_2$ begins when you interpret the slopes.  </div>



:::
::::


## 3D visualization? Out of Curiosity...

:::: {.columns}
::: {.column width="40%"}

The 3D visualization of a piecewise regression is actually kinda funny:


```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

library(FabioFun)

nice_3D_plot(y = Temp_avg_2021$Temp_F,
             x1 = Temp_avg_2021$D1,
             x2 = Temp_avg_2021$D2,
             axis_names = c("Temperature", 
                            "D1", 
                            "D2"),
             reg_plane = TRUE)
```


<div style="font-size: 24px"> You can tell that there actually is no "depth" to this figure, and all the dots are on the side of the box. The 3D representation just "folds" the 2D plot from some slides ago at the cut-off point! </div>

:::
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: false 
#| code-line-numbers: false

library(FabioFun)

nice_3D_plot(y = Temp_avg_2021$Temp_F,
             x1 = Temp_avg_2021$D1,
             x2 = Temp_avg_2021$D2,
             axis_names = c("Temperature", 
                            "D1", 
                            "D2"),
             reg_plane = TRUE)%>%  
                    bslib::card(full_screen = TRUE)



```


:::
::::

<div style="font-size: 18px; padding-top: 16px;"> The 2D visualization is clearly better, but just thought I would show this ðŸ¤· </div>

## Spline Regression

Spline regression is quite similar to loess regression. Spline regression is a piecewise regression with many segments, and is very similar to loess regression.

:::: {.columns}
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125


spline_fit <- lm(Temp_F ~ splines2::bsp(Temp_avg_2021$time_cent, df = 5), 
                 data = Temp_avg_2021)
summary(spline_fit)
```


:::
::: {.column width="30%"}

We can use the `bsp()` function form the `splines2` package to define our spline regression. the `df =` arguments sets the number of independent segments. here I choose 5 segments. 


The coefficients are not that helpful. Let's look at the visualization.

:::
::::


## Visulizing Spline Regression

The visualization for spline regression requires you calculate the predictions from the regression model (unless I decide to make a function at some point):

:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

pred_data <- seq(min(Temp_avg_2021$time_cent),
                 max(Temp_avg_2021$time_cent))

pred <- predict(spline_fit, 
                newdata = list(time_cent = pred_data))

# The code to generate the plot is in the drop down menu
```

<div style="font-size: 22px"> You can see that the spline regression line (red) is very similar to the loess regression line (blue). The two methods are quite similar. </div>

<div style="font-size: 22px; padding-top: 16px;"> Also note that the spline regression line is smooth. This is because by default the `slpine2` package uses 3<sup>rd</sup> degree polynomial segments. (see `help(bsp)`) </div>
 
 
:::
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: plot-code

ggplot(Temp_avg_2021,
       aes(x = time_cent, y = Temp_F)) +
        geom_point() +
  geom_smooth(method = "loess", 
              formula = y ~ x,
              se = FALSE) +
  geom_line(aes(x = pred_data,
                y = pred),
            col = "red",
            size = 1) 

```


:::
::::




## References 

<div id="refs"> </div>





