---
title: "Lab 11: Interactions With Categorical Predictors"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "Lab 11: Interactions With Categorical Predictors"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

# install.packages("tidyverse")
install.packages("emmeans")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(tidyverse)
theme_set(theme_classic(base_size = 16, 
                        base_family = 'serif'))
library(emmeans)
```


</br>

<div style="font-size: 26px">

::: {.panel-tabset}
### `emmeans`

<div style="font-size: 22px"> The `emmeans` package [@Lenth_etal_2025] is a great package for exploring the results of many linear, generalized linear, and mixed models. it also helps compute mean contrasts, trends, and comparisons of slopes. I discover some new functionality every time I look at this package!  </div>

:::

</div>

  

:::


::: {.column width="50%"}

<center style="padding-bottom: 41px;"> [Data]{.data-title} </center>

<div style="font-size: 22px"> We'll continue with the `hsb2` dataset from [Lab 10](https://quinix45.github.io/PSYC-7804-Regression-Lab-Slides/Slides%20Files/Lab%2010.html#/todays-packages-and-data){target="_blank"}. 
 </div>
 
 
```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

hsb2 <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/hsb2.csv")
```

<center style="padding-bottom: 11px;"> </center>

```{r}
reactable::reactable(hsb2,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1.105rem"),
                     pagination = FALSE, highlight = TRUE, height = 300)
```

:::
::::



## What are we looking at Today?

In Lab 10 we looked at categorical predictors and a bunch of coding schemes for categorical variables, but we will stick to [dummy coding](https://quinix45.github.io/PSYC-7804-Regression-Lab-Slides/Slides%20Files/Lab%2010.html#/creating-dummy-coded-columns){target="_blank"} for this lab.


::: {.fragment fragment-index=1}
We eventually want to look at whether `gender` ($Z$) moderates how `socst` ($X$) score predicts `write` ($Y$) score. As a reminder:
:::

:::: {.columns}
::: {.column width="50%"}

<ul style="font-size: 28px">  

::: {.fragment fragment-index=2}

<li> `gender`: student gender (either *male* or *female*).  </li>
<li> `socst`: score on a standardized social studies assessment. </li>
<li> `write`: score on a standardized writing assessment. </li>

:::

</ul>

::: {.fragment fragment-index=3}

<div style="font-size: 26px"> let's center `socst` for ease of interpretation later: </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

hsb2$socst_cent <- scale(hsb2$socst, scale = FALSE)[,1]
```
:::

:::


::: {.column width="50%"}

::: {.fragment fragment-index=4}
`gender` is a categorical variable, so let's turn it into a `factor`

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

hsb2$gender <- factor(hsb2$gender)

contrasts(hsb2$gender)
```
:::

:::
::::

::: {.fragment fragment-index=5}
<div style="text-align:center; padding-top: 20px; font-size: 30px;">
So for the dummy coded `gender` variable, $0 =$ *female* and $1 =$ *male* 
</div> 
:::

## Starting from The Top

First let's start with a regression with the `gender` only predicting `write` score. We saw this in Lab 10 already.
  

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_gender <- lm(write ~ gender, data = hsb2)

summary(reg_gender)
```

:::
::: {.column width="50%"}

$$\widehat{\mathrm{write}} = 54.99 - 4.86 \times \mathrm{gender} $$


::: {.fragment fragment-index=1}
We looked at the interpretation for this kind of regression in the last lab:
:::

<ul style="font-size: 26px">  

::: {.fragment fragment-index=1}
<li> $54.99$ is the expected mean `write` score of *females* (which is coded as $0$). </li>
:::

::: {.fragment fragment-index=2}
<li> $-4.86$ is the expected mean difference in `write` score for *males* compared to *females*. </li>
:::

</ul>

::: {.fragment fragment-index=3}
Thus, there is a significant difference in `write` score between the two groups, with *males* being $-4.86$ lower on average.  
:::

:::
::::

## Visualizing `write ~ gender`

:::: {.columns}
::: {.column width="30%"}

Once again, we can visualize the mean differences between the two groups.


This is nothing new, but it is interesting to see what happens to this plot once we add a continuous predictor to our regression.


:::
::: {.column width="70%"}



```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"


mean_female <- mean(hsb2$write[hsb2$gender == "female"])
mean_male <- mean(hsb2$write[hsb2$gender == "male"])

ggplot(hsb2, aes(x = gender, y = write)) +
  geom_point() +
  geom_hline(aes(yintercept = mean(mean_female), color = "Female"), 
             linetype = "dashed") +
  geom_hline(aes(yintercept = mean(mean_male), color = "Male"), 
             linetype = "dashed") +
  scale_color_manual(values = c("Female" = "blue", "Male" = "red")) +
  labs(color = "Means")
```


:::
::::

## Adding a continous Predictor

We further hypothesize that `socst` should be positively related to `write` score.

:::: {.columns}
::: {.column width="50%"}


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_both <- lm(write ~ gender + socst_cent, data = hsb2)

summary(reg_both)
```
:::

::: {.column width="50%"}

$$\widehat{\mathrm{write}} = 54.72 - 4.28 \times \mathrm{gender} +  .52 \times \mathrm{socst}$$

::: {.fragment fragment-index=1}
Now that we have a continuous predictor, things are slightly different:
:::

<ul style="font-size: 26px">  

::: {.fragment fragment-index=2}
<li> $54.72$ is now the expected mean `write` score of *females*, when `socst` is at its mean value. </li>
:::

::: {.fragment fragment-index=3}
<li> $-4.28$ is the expected mean difference in `write` score for *males* compared to *females*. when `socst` is controlled. </li>
:::

::: {.fragment fragment-index=4}
<li> $.52$ is the expected mean increase in `write` for each unit increase in `socst`. Unsurprisingly, `write` and `socst` are positively related. </li>
:::

</ul>

:::
::::


## Visualizing `write ~ gender + socst`


:::: {.columns}
::: {.column width="30%"}

<div style="font-size: 26px"> What this model represents is effectively two separate regression lines for the two groups, which differ only based on the intercept. </div>

</br>

::: {.fragment fragment-index=1}
<div style="font-size: 26px"> If you have heard of the term [ANCOVA](https://en.wikipedia.org/wiki/Analysis_of_covariance){target="_blank"} before, this is what we just ran.  </div>
:::


</br>

::: {.fragment fragment-index=2}
<div style="font-size: 26px"> Note that the slopes are the same, meaning that we are assuming that the relation between `socst` and `write`  should be the same for both groups. </div>
:::

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"


ggplot(hsb2, aes(x = socst_cent, y = write, colour = gender)) +
  geom_point(aes(shape = gender), alpha= .6) +
  geom_abline(intercept = coef(reg_both)[1], slope = coef(reg_both)[3], col = "purple") + 
  geom_abline(intercept = coef(reg_both)[1] + coef(reg_both)[2], slope = coef(reg_both)[3], col = "red") + 
  scale_color_manual(values=c( "purple", "red")) 
```
:::
::::

## Adding the Interaction

If we think that the relationship between `socst` and `write` should change depending on `gender`, we can test this hypothesis by including an interaction effect.

:::: {.columns}
::: {.column width="55%"}


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_int <- lm(write ~ gender*socst_cent, data = hsb2)

summary(reg_int)
```
:::

::: {.column width="45%"}


::: {.fragment fragment-index=1}
we add the $.2 \times \mathrm{gender} \times \mathrm{socst}$ term to the equation from two slides ago. 
:::

<ul style="font-size: 24px">  

::: {.fragment fragment-index=1}
<li> $54.78$ is still the expected mean `write` score of *females*, when `socst` is at its mean value. </li>
:::
::: {.fragment fragment-index=2}
<li> $-4.28$ is the expected mean difference in `write` score for *males* compared to *females*, when `socst` is controlled. </li>
:::

::: {.fragment fragment-index=3}
<li> $.42$ is the expected mean increase in `write` for each unit increase in `socst` *for females*. </li>
:::

::: {.fragment fragment-index=4}
<li> $.2$ is the expected *difference in slope* between `write` and `socst` for the *males*. </li>
:::

</ul>

:::
::::


## Interpreting the Interaction Term

When we have a categorical moderator, interpreting the interaction terms boils down at looking how the slopes of the groups change.

:::: {.columns}
::: {.column width="50%"}


<center> **Equation for Males** 


<div style="font-size:18px;"> 
$$\widehat{\mathrm{write}} = 54.78 - 4.28 \times \mathrm{gender} +  .42 \times \mathrm{socst} + .2 \times \mathrm{gender} \times \mathrm{socst}$$ 
</div>


::: {.fragment fragment-index=1}
&#8595;



<div style="font-size:18px;"> 
$$\widehat{\mathrm{write}} = 54.78 - 4.28 \times \mathrm{gender} +  .42 \times \mathrm{socst} + .2 \times 1 \times \mathrm{socst}$$ 
</div>
:::

::: {.fragment fragment-index=2}
&#8595;

<div style="font-size:18px;"> 
$$\widehat{\mathrm{write}} = 54.78 - 4.28 \times \mathrm{gender} +  .42 \times \mathrm{socst} + .2 \times \mathrm{socst}$$ 
</div>
:::

::: {.fragment fragment-index=3}
&#8595;

<div style="font-size:18px;"> 
$$\widehat{\mathrm{write}} = 54.78 - 4.28 \times \mathrm{gender} +  .62 \times \mathrm{socst}$$ 
</div>
:::

</center>

:::
::: {.column width="50%"}

<center> **Equation for Females**

<div style="font-size:18px;"> 
$$\widehat{\mathrm{write}} = 54.78 - 4.28 \times \mathrm{gender} +  .42 \times \mathrm{socst} + .2 \times \mathrm{gender} \times \mathrm{socst}$$ 


</div>


::: {.fragment fragment-index=1}
&#8595;

<div style="font-size:18px;"> 
$$\widehat{\mathrm{write}} = 54.78 - 4.28 \times \mathrm{gender} +  .42 \times \mathrm{socst} + .2 \times 0 \times \mathrm{socst}$$ 
</div>
:::


::: {.fragment fragment-index=2}
&#8595;

<div style="font-size:18px;"> 
$$\widehat{\mathrm{write}} = 54.78 - 4.28 \times \mathrm{gender} +  .42 \times \mathrm{socst}$$ 
</div>
:::

</center>

:::
::::

::: {.fragment fragment-index=3}
So, we expect the relation between `write` and `socst` (i.e., the *slope* of `socst`) to be $.2$ larger for `males`. 
:::


## Visualizing `write ~ gender * socst`

:::: {.columns}
::: {.column width="30%"}

</br>

<div style="font-size: 28px; padding-bottom:14px;"> The slope for `males` is steeper. Although at lower levels of `socst`, `females` do considerably better, the gap closes more and more as `socst` increases.  </div>

::: {.fragment fragment-index=1}
<div style="font-size: 28px"> A significant interaction term implies that the difference between the slopes between the two groups is non-zero. </div>
:::

:::
::: {.column width="70%"}

::: {.panel-tabset}

### `write ~ gender * socst`

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"


ggplot(hsb2, aes(x = socst_cent, y = write, colour = gender)) +
  geom_point(aes(shape = gender), alpha= .6) +
  geom_abline(intercept = coef(reg_int)[1], slope = coef(reg_int)[3], col = "purple") + 
  geom_abline(intercept = coef(reg_int)[1] + coef(reg_int)[2], slope = coef(reg_int)[3] + coef(reg_int)[4], col = "red") + 
  scale_color_manual(values=c( "purple", "red")) 
```

### `write ~ gender + socst`

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"


ggplot(hsb2, aes(x = socst_cent, y = write, colour = gender)) +
  geom_point(aes(shape = gender), alpha= .6) +
  geom_abline(intercept = coef(reg_both)[1], slope = coef(reg_both)[3], col = "purple") + 
  geom_abline(intercept = coef(reg_both)[1] + coef(reg_both)[2], slope = coef(reg_both)[3], col = "red") + 
  scale_color_manual(values=c( "purple", "red")) 
```


### `write ~ gender`

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"


mean_female <- mean(hsb2$write[hsb2$gender == "female"])
mean_male <- mean(hsb2$write[hsb2$gender == "male"])

ggplot(hsb2, aes(x = gender, y = write)) +
  geom_point() +
  geom_hline(aes(yintercept = mean(mean_female), color = "Female"), 
             linetype = "dashed") +
  geom_hline(aes(yintercept = mean(mean_male), color = "Male"), 
             linetype = "dashed") +
  scale_color_manual(values = c("Female" = "blue", "Male" = "red")) +
  labs(color = "Means")
```


:::

:::
::::


##  Simple Slopes With `emmeans`

Whenever categorical variables are involved, the `emmeans` package likely has some functions to summarize results. For interaction between continuous and categorical variables, the `emtrends()` function is very helpful!


:::: {.columns}
::: {.column width="40%"}

</br>

:::{.fragment fragment-index=1}
<div style="font-size: 26px"> Although we calculated the simple slopes by hand two slides ago, you don't really want to do that (especially when your categorical predictors have more levels!) </div>


<div style="font-size: 26px; padding-top: 20px;"> The `emtrends()` function also provides confidence intervals for the slopes. </div>
:::

:::
::: {.column width="60%"}

::: {.fragment fragment-index=2}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

emtrends(reg_int, ~gender, var = "socst_cent")
```

<ul style="font-size: 22px">  

<li>  `reg_int` is the  regression model. </li>

<li>  `~` followed by `gender` implies that we want the slopes for each level of the `gender` variable. </li>

<li>  `var = ` followed by `"socst_cent"` is the continuous predictor we want to get slopes for. </li>


</ul>
:::


:::
::::

::: {.fragment fragment-index=3}
<div style="font-size: 24px"> Here we have a simple example, but you can request slopes for multiple predictors and categorical moderators at once. [This page](https://stats.oarc.ucla.edu/r/seminars/interactions-r/){target="_blank"} has a really comprehensive tutorial on how to use the `emmeans` package with interactions. 
 </div>
:::


## Predicted Means 

<div style="font-size: 24px"> So far we have been focusing on the slope of `socst_cent`, but we can also shift our focus to means and mean differences. In our case, the `emmeans` package can be used to calculate predicted means of groups *at different values* of the continuous variable. </div>

:::: {.columns}
::: {.column width="40%"}

::: {.fragment fragment-index=1}
<div style="font-size: 24px; padding-top: 12px;"> We need to define the values of `socst_cent` at which we want to calculate means. We do this by creating a `list` object containing the name of the continuous variable and the values at which we want to calculate means. </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

sd_below <- -sd(hsb2$socst_cent)
sd_above <- sd(hsb2$socst_cent)

# the mean is 0 because the variable is centered
mylist <- list(socst_cent=c(sd_below,
                            0,
                            sd_above))
```

:::
:::

::: {.column width="60%"}

::: {.fragment fragment-index=2}

<div style="font-size: 24px; padding-top: 12px;"> Then we pass the regression object (`reg_int`), and use `~socst_cent*gender` to tell the function that we want the means of each level of `gender` by the specified values of `socst_cent`: </div>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

means <- emmeans(reg_int, ~socst_cent*gender, 
                at = mylist)
```
```{r}
means
```


:::

:::
::::


::: {.fragment fragment-index=3}
<div style="font-size: 26px; padding-top: 12px;"> In the **emmean** (stands for *estimated marginal mean*) column we have the mean expected value of `write` score for males and females at -/+ 1 SD and mean of `socst` score. </div>
:::

## Predicted Means Graphically

:::: {.columns}
::: {.column width="40%"}

The previous slide may seem intimidating, but the means of `write` score that we just calculated (brown dots in the plot) are simply the values of $y$-axis given some values of $x$ according to the regression line of each groups.

```{r}
means
```

:::
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"


ggplot(hsb2, aes(x = socst_cent, y = write, colour = gender)) +
  geom_point(aes(shape = gender), alpha= .6) +
  geom_abline(intercept = coef(reg_int)[1], slope = coef(reg_int)[3], col = "purple") + 
  geom_abline(intercept = coef(reg_int)[1] + coef(reg_int)[2], slope = coef(reg_int)[3] + coef(reg_int)[4], col = "red") + 
  scale_color_manual(values=c( "purple", "red")) +
  geom_point(aes(x = -10.7, y = 50.3), color = "brown", size = 4) +
  geom_point(aes(x = -10.7, y =  43.8), color = "brown", size = 4) +
  geom_point(aes(x = 0, y = 54.8), color = "brown", size = 4) +
  geom_point(aes(x = 0, y =  50.5), color = "brown", size = 4) +
  geom_point(aes(x = 10.7, y = 59.3), color = "brown", size = 4) +
  geom_point(aes(x = 10.7, y = 57.2), color = "brown", size = 4) 
  
```

:::
::::


##  Testing Mean Differences

<div style="font-size: 24px"> Now that we have the means of `write` score for both groups at different levels of `socst` score, the logical extention is to test whether these means are significantly different from each other. </div>


:::: {.columns}
::: {.column width="60%"}

::: {.fragment fragment-index=1}

We use the `contrast()` function from the `emmeans` package (careful! not the `contrasts()` function 🙈):

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

contrast(means, "pairwise", by= "socst_cent")
```

:::
:::

::: {.column width="40%"}

::: {.fragment fragment-index=2}

<div style="font-size: 26px; padding-top: 12px;"> As we observed previously, the mean difference in `write` is significant at lower values of `socst`, but is no longer significant at high values of `socst`. </div>
:::


::: {.fragment fragment-index=3}
<div style="font-size: 26px; padding-top: 12px;"> Because we are running multiple comparisons, there is something to be said for adjusting the $p$-values for Type I error rates, although that takes a bit more work and I don't really want to spend time on that (and you know [how I feel about type I errors](https://statmodeling.stat.columbia.edu/2004/12/29/type_1_type_2_t/){target="_blank"} 🤷). </div>
:::

:::
::::

## Mean differences Graphically

:::: {.columns}
::: {.column width="35%"}

<div style="font-size: 24px;"> Testing mean differences at different values of `socst` simply mean testing whether the dashed black lines are significantly different from 0 in length. The dashed lines, representing the distance between the two regression lines becomes smaller and smaller as `socst` increases.  </div>

::: {.fragment fragment-index=1}
::: {.callout-note}

## It's a matter of perspective

You may have noticed that first we looked at *(1)* whether the slope of `socst` changed depending on `gender`, and then we *(2)* looked at whether there were mean differences in mean `write` score for `gender` depending on different values of `socst`. In other words, we swapped the variable that we were treating as the moderator. This highlights how from a statistical perspective there is no difference between a moderator and a focal predictor. You decide how two present the results and frame the two variables.

::: 
:::

:::
::: {.column width="65%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"

# save summary for mean differences
sum <- summary(means)

ggplot(hsb2, aes(x = socst_cent, y = write, colour = gender)) +
  geom_point(aes(shape = gender), alpha= .2) +
  geom_abline(intercept = coef(reg_int)[1], slope = coef(reg_int)[3], col = "purple") + 
  geom_abline(intercept = coef(reg_int)[1] + coef(reg_int)[2], slope = coef(reg_int)[3] + coef(reg_int)[4], col = "red") + 
  scale_color_manual(values=c( "purple", "red")) +
  geom_segment(y = sum[1,3], yend = sum[4,3], x = sum[4,1], xend = sum[4,1], col = "black", lty = 2) +
  geom_segment(y = sum[2,3], yend = sum[5,3], x = sum[5,1], xend = sum[5,1], col = "black", lty = 2) +
  geom_segment(y = sum[6,3], yend = sum[3,3], x = sum[3,1], xend = sum[3,1], col = "black", lty = 2) 
  
```


:::
::::

## johnson-neyman Plot?

:::: {.columns}
::: {.column width="35%"}



<div style="font-size: 24px;"> Instead of testing differences in the slopes for the 2 groups at different values of the continuous variable, we can do that for all possible values of the continuous variable by using a Johnson-Neyman plot </div>

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
#| classes: code-125

# the predictor must be a vector of 0s and 1s
hsb2$gender_bin <- ifelse( hsb2$gender == "female", 0, 1)

# rerun the regression with binary gender
reg_int_2 <- lm(write ~ gender_bin * socst_cent, data = hsb2)

# save summary for mean differences
interactions::johnson_neyman(reg_int_2, 
              pred = "gender_bin",
              modx = "socst_cent")
```

<div style="font-size: 24px;"> Because we have a categorical variable, the interpretation is conceptually different than what we saw in [Lab 9](https://quinix45.github.io/PSYC-7804-Regression-Lab-Slides/Slides%20Files/Lab%209.html#/the-johnson-neyman-plot){target="_blank"} </div>
 

:::
::: {.column width="65%"}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125


# the predictor must be a vector of 0s and 

hsb2$gender_bin <- ifelse( hsb2$gender == "female", 0, 1)

reg_int_2 <- lm(write ~ gender_bin * socst_cent, data = hsb2)

# save summary for mean differences
interactions::johnson_neyman(reg_int_2, 
                             pred = "gender_bin",
                             modx = "socst_cent")$plot
```


<div style="font-size: 24px;"> The $y$-axis is the difference between the two group means (i.e., the *distance* between the two lines on the previous slide) in `write` score for each value of `socst` score. Consistent with the plot on the previous slide, the difference is no longer significant when `socst` is around 10 or more. </div>

:::

:::
::::



## Plotting with `emmip()`

We can use the `emmip()` to plot the slopes for both groups. 

:::: {.columns}
::: {.column width="40%"}

::: {.fragment fragment-index=1}
<div style="font-size: 23px"> First, we need to create a list that tells `emmip()` the range of the continuous variable for which want to plot slopes. here I choose the minimum and maximum of `socst_cent`. </div>

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| classes: code-125

min <- min(hsb2$socst_cent)
max <- max(hsb2$socst_cent)

mylist <- list(socst_cent = c(min, 
                              max))
```
:::

::: {.fragment fragment-index=2}
<div style="font-size:23px; padding-top: 12px;"> Then we pass pass the regression object, the categorical variable by which we want the slopes, and the list containing the range of the x-axis. </div>

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
#| classes: code-125

emmip(reg_int, gender~socst_cent, at = mylist)
```
:::

:::
::: {.column width="60%"}

::: {.fragment fragment-index=2}
```{r}
emmip(reg_int, gender~socst_cent, at = mylist)
```
:::
:::
::::

::: {.fragment fragment-index=3}
<div style="font-size: 23px; padding-top: 14px;"> This may seem a bit complicated for no reason 🤨 but it affords lots of flexibility when your model is more complex than what we have here. </div>

<div style="font-size: 16px"> I personally prefer plotting things "manually" with `ggplot` 🫣, although I understand that may not be everyone's cup of tea </div>
:::


## Plotting with `interact_plot()`

In our case, it is slightly simpler to use the `interact_plot()` function, but, unlike `emmip()`, it requires a binary variable and only works with 2 categories.

:::: {.columns}
::: {.column width="40%"}



```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
#| classes: code-125

interactions::interact_plot(reg_int_2, 
              modx = "gender_bin",
              pred = "socst_cent",
              modx.values = c(0, 1))
```

Notice that I am using the regression model with the binary version of `gender` because the `interactions` package does not play nice factor variables.

::: {.fragment fragment-index=1}

So, you have multiple options for plotting simple slopes. Still, my go to is just using `ggplot` because it gives me the most freedom (as you can tell from earlier slides).   

:::



:::
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

interactions::interact_plot(reg_int_2, 
              modx = "gender_bin",
              pred = "socst_cent",
              modx.values = c(0, 1))
```

:::
::::





## References 

<div id="refs"> </div>







