---
title: "Lab 4: Two-Predictor Regression Part 1"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
format:
   revealjs:
      footer: "Lab 4: Two-Predictor Regression Part 1"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---

## Today's Packages and Data ðŸ¤—

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150


install.packages("GGally")
install.packages("plotly")
install.packages("rio")
install.packages("tidyverse")

```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(GGally)
library(plotly)
library(rio)
library(tidyverse)
theme_set(theme_classic(base_size = 14, 
                        base_family = 'serif'))
```

</br>

<div style="font-size: 26px">

::: {.panel-tabset}

### `GGally`

The ``GGally` package [@Schloerke_etal_2024] builds upon `ggplot2` and includes many fucntions for creating complex plots. 

### `plotly`

The `plotly` package [@Sievert_etal_2024a] is a Python and R package used to create interactive visualizations with JavaScript elements. 

### `rio`

The `rio` package [@Becker_etal_2024] developers describe this package as the *Swiss-Army Knife for Data I/O*. The `import()` and `export()` functions can import/export just about any data type.

### `tidyverse`

The `tidyverse` package [@Wickham_RStudio_2023] loads a suite of packages that help with data cleaning and visualization. Among others, `tidyverse` loads both `dplyr` and `ggplot2`.


:::
</div>

:::
::: {.column width="50%"}

<ul>

<li> Let's also load the [data](https://worldhappiness.report/ed/2024/){target="_blank"}
 for today: </li>
</ul>

```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

WH_2024 <- import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/World_happiness_2024.csv")

# let's peak at our variables
str(WH_2024, vec.len = 2)
```

:::
::::


## Variables of Interest

We eventually want to look at how `Log_GDP` and `Freedom` impact `Happiness_score` across the countries in our data. Let's first explore the variables.


:::: {.columns}
::: {.column width="50%"}

<ul style="font-size: 24px">  

<li>  The corraltion matrix for the variables: </li>

</ul>


```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# selecting just the variables of interest makes code cleaner later
reg_vars <- WH_2024[, c("Happiness_score",
                        "Log_GDP",
                        "Freedom")]

# Just the correlation table
cor(reg_vars)
```

:::
::: {.column width="50%"}

<ul style="font-size: 24px">  

<li>  Some descriptive statistics: </li>

</ul>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

psych::describe(reg_vars)
```

:::
::::

## Helpful Visualization

The `ggpairs()` function from the `GGally` package creates a visualizations that provides a lot of information in one go!

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| out-width: "70%"


ggpairs(reg_vars)
```

## Individual Regressions


:::: {.columns}
::: {.column width="50%"}

<ul style="font-size: 24px">  

<li>  `Log_GDP` only regression </li>

</ul>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_GDP <- lm(Happiness_score ~ Log_GDP, 
              reg_vars)
summary(reg_GDP)
```
:::
::: {.column width="50%"}

<ul style="font-size: 24px">  

<li>  `Freedom` only regression </li>

</ul>

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_free <- lm(Happiness_score ~ Freedom, 
              reg_vars)
summary(reg_free)
```
:::
::::

## Regression With Both Predictors


:::: {.columns}
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_full <- lm(Happiness_score ~ Log_GDP + Freedom, 
          reg_vars)

summary(reg_full)
```

:::
::: {.column width="40%"}

As you can see, the regression coefficients of the two predictors change when you both variables are included ðŸ§

<br>

Let's visualize what is happening.

:::
::::



## Individual Regression plots 

These are the equivalent plots to the individual regressions:

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

  ggplot(reg_vars,  
 aes(x = Log_GDP, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm", 
                formula = "y~x", 
                se = FALSE)
```



:::
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"

  ggplot(reg_vars,  
 aes(x = Freedom, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm", 
                formula = "y~x", 
                se = FALSE)
```

:::
::::


## Adding the line from multiple regression? 

Let's add the regression lines estimated from the regression with both predictors...

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"


  ggplot(reg_vars,  
 aes(x = Log_GDP, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm", 
                formula = "y~x", 
                se = FALSE) +
    geom_abline(intercept = coef(reg_full)[1], 
                slope = coef(reg_full)[2])
```


:::
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot Code"


  ggplot(reg_vars,  
 aes(x = Freedom, y = Happiness_score)) +
 geom_point() + 
    geom_smooth(method = "lm", 
                formula = "y~x", 
                se = FALSE) +
    geom_abline(intercept = coef(reg_full)[1], 
                slope = coef(reg_full)[3])
```
:::
::::

Hold up, the lines from the multiple regression results seem way off !?

ðŸ’¡ Maybe we need a change of perspective!


## Looking Inside the Box

Now that we are dealing with 3 variables (`Log_GDP`, `Freedom`, `Happiness_score`), our data points are in a 3D box, not a 2D plot. 

:::: {.columns}
::: {.column width="40%"}

The 2D plots from the previous slide are just the *sides of the box* ðŸ¤“

I made a function for interactive 3D plot. It has a bunch of useful features that I will show later

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

source("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/R_scripts/3D_plot.R")

nice_3D_plot(y = reg_vars$Happiness_score,
             x1 = reg_vars$Log_GDP,
             x2 = reg_vars$Freedom,
             dot_labels = WH_2024$Country_name,
             axis_names = c("Happiness", 
                            "GDP", 
                            "Freedom"))
```

:::
::: {.column width="60%"}

```{r out.width = "100%"}
source("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/R_scripts/3D_plot.R")


nice_3D_plot(y = reg_vars$Happiness_score,
                            x1 = reg_vars$Log_GDP,
                            x2 = reg_vars$Freedom,
                            dot_labels = WH_2024$Country_name,
                              axis_names = c("Happiness", 
                                             "GDP", 
                                             "Freedom"),
                              plane_res = 20,
                              reg_plane = FALSE)%>%  
                    bslib::card(full_screen = TRUE)
```


:::
::::

## The Regression Plane 

<div style="font-size: 24px">  

Remember that in linear regression with 1 predictor we are looking for the line that on average is closest to all point. When we have 2 predictors we are looking for the **plane** that is closest to all the points!

</div>



:::: {.columns}
::: {.column width="40%"}

```{r}
#| eval: false
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = reg_vars$Happiness_score,
             x1 = reg_vars$Log_GDP,
             x2 = reg_vars$Freedom,
             dot_labels = WH_2024$Country_name,
             axis_names = c("Happiness", 
                            "GDP", 
                            "Freedom"),
             plane_res = 20,
             reg_plane = TRUE) 

```


:::
::: {.column width="60%"}

```{r out.width = "100%"}
#| eval: true
#| echo: false
#| code-line-numbers: false
#| classes: code-125

nice_3D_plot(y = reg_vars$Happiness_score,
             x1 = reg_vars$Log_GDP,
             x2 = reg_vars$Freedom,
             dot_labels = WH_2024$Country_name,
             axis_names = c("Happiness", 
                            "GDP", 
                            "Freedom"),
             plane_res = 20,
             reg_plane = TRUE) %>%  
  bslib::card(full_screen = TRUE)

```

:::
::::

## Back to our friend *R* <sup>2</sup>

$R^2$ is often referred to as the *variance explained* in the dependent variable. The term "variance explained" made no sense to me the first time I heard it ðŸ¤·






## References 

<div id="refs"> </div>


