---
title: "Lab 6: Semi-partial, Partial-correlations, and Model Comparison"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "Lab 6: Variance Explained, Semi-partial and Partial-correlations"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---

## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

install.packages("GGally")
install.packages("plotly")
install.packages("rio")
install.packages("tidyverse")

```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(GGally)
library(plotly)
library(rio)
library(tidyverse)
theme_set(theme_classic(base_size = 14, 
                        base_family = 'serif'))
```

</br>

<div style="font-size: 26px">

::: {.panel-tabset}

### `GGally`

The ``GGally` package [@Schloerke_etal_2024] builds upon `ggplot2` and includes many fucntions for creating complex plots. 

### `plotly`

The `plotly` package [@Sievert_etal_2024a] is a Python and R package used to create interactive visualizations with JavaScript elements. 

### `rio`

The `rio` package [@Becker_etal_2024] developers describe this package as the *Swiss-Army Knife for Data I/O*. The `import()` and `export()` functions can import/export just about any data type.

### `tidyverse`

The `tidyverse` package [@Wickham_RStudio_2023] loads a suite of packages that help with data cleaning and visualization. Among others, `tidyverse` loads both `dplyr` and `ggplot2`.


:::
</div>

:::
::: {.column width="50%"}

<ul>

<li> Let's also load the [data](https://worldhappiness.report/ed/2024/){target="_blank"}
 for today: </li>
</ul>

```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

# Data from https://library.virginia.edu/data/articles/hierarchical-linear-regression

# myData <- rio::import('http://static.lib.virginia.edu/statlab/materials/data/hierarchicalRegressionData.csv')


# let's peak at our variables
str(WH_2024, vec.len = 2)
```

:::
::::






## References 

<div id="refs"> </div>

