---
title: "Lab 10: Categorical Predictors"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
notice: |
  @Wickham_RStudio_2023
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"

format:
   revealjs:
      footer: "Lab 10: Categorical Predictors"
      width: 1280
      height: 720
      chalkboard: true
      revealjs-fontawesome: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

  
editor: source
---

## Font-awesome slide{visibility="hidden"}

uncounted and not visible slide to trigger the fontawesome extension

{{< fa thumbs-up >}} 


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

# install.packages("tidyverse")
install.packages("lubridate")
install.packages("splines2")
```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(tidyverse)
theme_set(theme_classic(base_size = 14, 
                        base_family = 'serif'))
library(lubridate)
library(splines2)
```


</br>

<div style="font-size: 26px">

::: {.panel-tabset}
### `lubridate`

The `lubridate` package [@Spinu_etal_2024] provides many functions to help you work with date and times.

### `splines2`

The `splines2` package [@Wang_etal_2024] can be used to calculate splines. 

:::

</div>


:::


::: {.column width="50%"}

<center style="padding-bottom: 41px;"> [Data]{.data-title} </center>

<div style="font-size: 22px"> This data come from a random sample of 200 students from the 1982 [high-school and beyond longitudinal study](https://nces.ed.gov/surveys/hsb/){target="_blank"}. See [here](https://www.openintro.org/data/index.php?data=hsb2){target="_blank"} to find out more about the variables. 
 </div>



```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

hsb2 <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/hsb2.csv")
```

<center style="padding-bottom: 11px;"> </center>

```{r}
reactable::reactable(hsb2,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1.105rem"),
                     pagination = FALSE, highlight = TRUE, height = 300)
```


:::
::::


## Categorical Variables in Regression

So far we have only dealt with continuous variables. Namely, we have interpreted slopes as "the change in $Y$ per unit change in $X$". 

<div style="font-size: 24px"> This interpretation only makes sense if both $Y$ and $X$ are continuous. But if we want to say that in this data `gender` should predict `write` score, the "unit-change" notion does not make as much sense. </div>




:::: {.columns}
::: {.column width="40%"}

<div style="font-size: 24px; padding-top: 12px;"> If we try to plot the `write` score on the $y$-axis and `gender` on the $x$-axis we can visualize observations in each group. But there are no "units" on the $x$-axis 🤔 </div>


<div style="font-size: 26px; padding-top: 22px;"> However, it turns out that there are ways to trick regression into treating categorical variables as continuous! </div>


<div style="font-size: 26px; padding-top: 12px;"> But first, I would like to point something out... </div>


:::
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"

ggplot(hsb2, aes(x = gender, y = write)) +
  geom_point()
```




:::
::::

## Intercept only Regression

You can use the `lm()` function with only one variable. Meaning, you can run a regression without any predictors (**!?**). for example, for the `write` variable:

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

car::S(lm(write ~ 1, data = hsb2))
```

:::
::: {.column width="50%"}

I would like to draw attention to the intercept, $52.78$ and the residual *SD*, $9.47$. These values are the **mean** and ***SD*** of the `write` variable. 

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

mean(hsb2$write)
sd(hsb2$write)

```

<div style="font-size: 20px"> This is consistent with the representation of regression shown in the [appendix of Lab 2](https://quinix45.github.io/PSYC-7804-Regression-Lab-Slides/Slides%20Files/Lab%202.html#/the-true-regression-model){target="_blank"}, Here $Y \sim N(\mu = 52.78, \sigma = 9.47)$
 </div>


:::
::::

<div style="font-size: 24px; padding-top: 14px"> So regression with only the intercept estimates the mean and *SD* of $Y$. But what now? Let's look at the graphical representation of the intercept only model. </div>


## Intercept only model

If we only use the intercept in out model, the "regression line" will be flat and intercept the $y$-axis at the mean of $Y$.

:::: {.columns}
::: {.column width="30%"}

...but wait a moment, we have two groups, `female` and `male`. How about we give each group their own intercept? 

- this simply implies that we believe that the mean of the $Y$ variable, `write`, should differ based on `gender`. 

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"

ggplot(hsb2, aes(x = gender, y = write)) +
  geom_point() +
  geom_hline(yintercept = mean(hsb2$write))
```

:::
::::

## Group Means

In the previous plot we were using the "grand mean" of `write`. If we use the two means instead...

:::: {.columns}
::: {.column width="30%"}

We see that the means of the two groups are different.

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

hsb2 %>% 
  group_by(gender) %>% 
   summarise(Group_means = mean(write))
```

We can run a regression model that is the representation of the graph on the right

:::
::: {.column width="70%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125
#| code-fold: true
#| code-summary: "Plot code"


mean_female <- mean(hsb2$write[hsb2$gender == "female"])
mean_male <- mean(hsb2$write[hsb2$gender == "male"])

ggplot(hsb2, aes(x = gender, y = write)) +
  geom_point() +
  geom_hline(aes(yintercept = mean(mean_female), color = "Female"), 
             linetype = "dashed") +
  geom_hline(aes(yintercept = mean(mean_male), color = "Male"), 
             linetype = "dashed") +
    geom_hline(aes(yintercept = mean(hsb2$write), color = "Grand Mean")) +
  scale_color_manual(values = c("Female" = "blue", "Male" = "red", "Grand Mean" = "black")) +
  labs(color = "Means")
```

:::
::::

## Coding Categorical Variables

We can code our categorical variables such that they are treated as continuous variables. Usually, we treat one group as "0", and the other group as "1". Let's say that `male` is $0$ and `female` is $1$:

:::: {.columns}
::: {.column width="60%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

hsb2$gender_binary <- ifelse(hsb2$gender == "female", 1, 0)

reg_gender_bin <- lm(write ~ gender_binary, data = hsb2)
summary(reg_gender_bin)
```


:::
::: {.column width="40%"}

$\mathrm{write} = 50.12 + 4.87 \times \mathrm{gender}$


<div style="font-size: 22px"> Remember that in regression, the intercept is *expected mean value* of $Y$ when $X = 0$, and the slope is the *expected mean value increase* per unit increase in $X$. </div>


<ul style="font-size: 22px">  

<li> Then, since we coded $0$ to represent `male`, the intercept *is the mean of* the `male` group in `write` score.    </li>

<li> And, since we coded $1$ to represent `female`, the slope *is the difference in means* between the `male` and `female` group in `write` score ($50.12 + 4.87 = 54.99$).    </li>

</ul>





:::
::::


## Mean differences? Sounds familiar?

We just tested whether `male` and `female` are significantly different in mean `write` score. This is what you should know as an independent-samples *t*-test (**!**).

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

t.test(hsb2$write ~ hsb2$gender,
       var.equal = TRUE)

```

:::
::: {.column width="50%"}

You should see that *t*-values, degrees of freedom, and *p*-values are the same as those for the slope on the previous slide. 


::: {.callout-note}
## No Coincidences in Statistics

There are no coincidences in statistics. If two methods give some of the same results, they must be related in some way. When you see such patterns, ask yourself *why* two methods produce the same results. Once you answer that question, you will have gained tremendous insight.

:::

:::
::::

*t*-tests don't exist really. For some reason everyone decides to teach *t*-tests separately from regression, but they are simply a specific case of a regression.






## References 

<div id="refs"> </div>


# Apendix: Alternative Cohen's d caluclatations





