---
title: "Lab 6: Semi-partial, Partial-correlations, and Model Comparison"
author: "Fabio Setti"
institute: "PSYC 7804 - Regression with Lab"
bibliography: Additional files/R packages.bib
csl: Additional files/apa.csl
notice: |
  @Wickham_RStudio_2023
title-slide-attributes:
  data-transition: "zoom"
  data-visibility: "uncounted"
format:
   revealjs:
      footer: "Lab 6: Semi-Partial, Partial-Correlations, and Model Comparison"
      width: 1280
      height: 720
      chalkboard: true
      slide-number: c/t 
      theme: Fabio_theme/Fabio_theme.scss
      navigation-mode: linear
      controls: false
      auto-stretch: false
      header-includes:
        - <script src="Fabio_theme/Fabio_theme.js"></script>

editor: source
---

## Font-awesome slide{visibility="hidden"}

uncounted and not visible slide to trigger the fontawesome extension

{{< fa thumbs-up >}} 


## Today's Packages and Data 🤗

:::: {.columns}
::: {.column width="50%"}

```{r}
#| code-fold: true
#| eval: false
#| echo: true
#| code-line-numbers: false
#| code-summary: "Install Packages Code"
#| classes: code-150

install.packages("ppcor")
# install.packages("tidyverse")

```

```{r}
#| eval: true
#| echo: true
#| code-line-numbers: false
#| warning: false
#| classes: code-150

library(ppcor)
library(tidyverse)
theme_set(theme_classic(base_size = 16, 
                        base_family = 'serif'))
```

</br>

<div style="font-size: 26px">

::: {.panel-tabset}

### `ppcor`

The `ppcor` package [@Kim_2015] contains functions to calculate partial and semi-partial correlations.
:::

</div>

:::
::: {.column width="50%"}

<center> **Data** </center>

</br>

```{r}
#| warning: false
#| classes: code-125
#| echo: true
#| code-line-numbers: false
#| output: true

# Data from https://library.virginia.edu/data/articles/hierarchical-linear-regression

dat <- rio::import("https://github.com/quinix45/PSYC-7804-Regression-Lab-Slides/raw/refs/heads/main/Slides%20Files/Data/Hierarc_data.csv")
```

```{r}
reactable::reactable(dat,
                     style = list(fontFamily = "Work Sans, sans-serif", fontSize = "1.175rem"),
                     pagination = FALSE, highlight = TRUE, height = 350)
```


:::
::::


## Our Predictors

We  want to see how happiness (`happiness`) relates to age (`age`), and number of friends (`friends`).

:::: {.columns}
::: {.column width="60%"}

We can calculate the correlation among our predictors

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

cor(dat[,1:3])
```

:::
::: {.column width="40%"}

::: {.fragment fragment-index=1}

Although useful, correlation by itself may not be the measure we are interested in. Rather, we may want to know the relationship between two variables *after* accounting for another set of variables.

:::

:::
::::

<br>

::: {.fragment fragment-index=2}
**Partial correlation** and **semi-partial correlation** are what we are looking for in this case. 
:::


## Some Formulas for Reference

If $r_{xy}$ represent the [correlation](https://raw.githack.com/quinix45/PSYC-7804-Regression-Lab-Slides/main/Slides%20Files/Lab%202.html#/correlation){target="_blank"} between $x$ and $y$, then 


::: {.fragment fragment-index=1}
* **Partial correlation**: 

  * $r_{xy.z} = \frac{r_{xy} - r_{xz} \times r_{yz}}{\sqrt{1 - r_{xz}^2}\times\sqrt{1 - r_{yz}^2}}$, where $z$ is controlled for both $x$ and $y$
:::


::: {.fragment fragment-index=2}
* **Semi-partial correlation**: In semi-partial correlation, the $z$ variable is held constant only for 1 of the other 2 variables. That means that in our case, there are 2 semi-partial correlations that can be calculated
:::

:::: {.columns}
::: {.column width="50%"}


::: {.fragment fragment-index=3}
  * $r_{x(y.z)} = \frac{r_{xy} - r_{xz} \times r_{yz}}{\sqrt{1 - r_{yz}^2}}$, 
  
  where $z$ is controlled only for $y$
:::


:::
::: {.column width="50%"}

::: {.fragment fragment-index=4}
  * $r_{y(x.z)} = \frac{r_{xy} - r_{xz} \times r_{yz}}{\sqrt{1 - r_{xz}^2}}$, 
  
  where $z$ is controlled only for $x$
:::

:::
::::

::: {.fragment fragment-index=5}
::: {.callout-note}

## spot differences and similarities in related formulas

Notice that the numerator (top part) of these formulas is always the same, and the denominator (bottom part) changes slightly. This means that if you want to understand the differences, you should not look at the numerator at all (pretend it does not exist!), but only at the denominator. 

:::
:::

## Partial Correlation "by hand"

As always, I like calculating things by hand first to make sure I know what is going on 🤷 Let's say that $y =$ `happiness`, $x =$ `age`, and $z =$ `friends`. 

:::: {.columns}
::: {.column width="50%"}

::: {.fragment fragment-index=1}
Let's first get our vanilla correlations

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

r_xy <- cor(dat$age, dat$happiness)
r_xz <- cor(dat$age, dat$friends)
r_yz <- cor(dat$happiness, dat$friends)
```
:::

::: {.fragment fragment-index=2}

Then, the partial correlation between $x$ and $y$, accounting for $z$ is

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

(r_xy - r_xz*r_yz)/(sqrt(1 - r_xz^2) * sqrt(1 - r_yz^2))

```
:::

:::
::: {.column width="50%"}

::: {.fragment fragment-index=3}
Note that this is also equivalent to calculating the correlation between the residuals of $y$ and $x$ after a regression with $z$ as a predictor

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

resid_y_z <- residuals(lm(happiness ~ friends, dat))
resid_x_z <- residuals(lm(age ~ friends, dat))

cor(resid_x_z, resid_y_z)
```
:::

:::
::::

::: {.fragment fragment-index=4}

So, $r_{xy.z}-.16$ is the correlation between `happiness` and `age` after taking out the proportion of variance explained by `friends` in both variables. 

:::

## Semi-Partial Correlation "by hand"

The semi-partial correlations between $x$ and $y$ accounting for $z$ can either be:

:::: {.columns}
::: {.column width="50%"}

::: {.fragment fragment-index=1}

With the formulas from the previous slides:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

(r_xy - r_xz*r_yz)/(sqrt(1 - r_xz^2))
```
or 

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

(r_xy - r_xz*r_yz)/(sqrt(1 - r_yz^2))
```
:::

:::
::: {.column width="50%"}

::: {.fragment fragment-index=2}


Similarly, we can get this by using regression residuals:

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

cor(dat$happiness, resid_x_z)
```
and

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

cor(dat$age, resid_y_z)
```
:::

:::
::::


::: {.fragment fragment-index=3}

- $r_{y(x.z)} = -.15$ represents the relationship between $y$ and $x$ after the taking out of $x$ the variance explained by $z$.  

- $r_{x(y.z)} = -.16$ represents the relationship between $x$ and $y$ after the taking out of $y$ the variance explained by $z$.  

:::

::: {.fragment fragment-index=4}

<center> Which type of correlation to use? **it depends on your research question** </center>

:::

## Using `ppcor` Functions

In practice, we use the `ppcor` package 

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# partial correlation

partial_cor <- pcor(dat[,1:3])
partial_cor$estimate
```


:::
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# semi-partial correlation

semipartial_cor <- spcor(dat[,1:3])
semipartial_cor$estimate
```
<div style="font-size: 20px">  Careful! in the semi-partial correlation output, the variable in the row is the variable that has no variance taken out.</div>
:::
::::

::: {.fragment fragment-index=1}

The functions from `ppcor` work in the same way as the ones used in [Lab 2](https://raw.githack.com/quinix45/PSYC-7804-Regression-Lab-Slides/main/Slides%20Files/Lab%202.html#/corr_test-output){target="_blank"}, where significance and other information is saved to a separate element.

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

partial_cor$p.value
```


:::
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

semipartial_cor$p.value
```

:::
:::
::::


## Model Comparison: Gain in Prediction

We are often (actually always) interested in comparing different models. One often relevant question is: "is it *worth it* to add more variables as predictors?" 

::: {.fragment fragment-index=1}

In the case of regression, the most popular way of comparing models is by comparing $R^2$ from [nested models](https://www.statology.org/nested-model/){target="_blank"}. Let's go back to how happiness (`happiness`) relates to age (`age`); then, we want to know whether adding number of friends (`friends`) improves our prediction.

:::

:::: {.columns}
::: {.column width="50%"}


::: {.fragment fragment-index=2}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_age <- lm(happiness ~ age, dat)

summary(reg_age)$r.squared
```
So `age` roughly explains $3\%$ of the variance in `happiness`
:::


:::
::: {.column width="50%"}

::: {.fragment fragment-index=3}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

reg_age_fr <- lm(happiness ~ age + friends, dat)

summary(reg_age_fr)$r.squared
```
`age` and `friends` together roughly explains $13\%$ of the variance in `happiness`

:::

:::
::::

::: {.fragment fragment-index=4}
The answer seems fairly straightforward: the model with both `age` and `friends` does better. But there is a catch with $R^2$ 🤨 
:::

## The catch With $R^2$

The "problem" with most popular measures of model fit such as $R^2$ is that they always increase as variables are added, even when the added variables are unrelated to $Y$. 

:::: {.columns}
::: {.column width="70%"}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

# generate random variable
set.seed(8879)
dat$random_var <- rnorm(nrow(dat))

reg_age_fr_rand <- lm(happiness ~ age + friends + random_var, dat)

# before
summary(reg_age_fr)$r.squared

# after 
summary(reg_age_fr_rand)$r.squared 
```
:::

:::
::: {.column width="30%"}

::: {.fragment fragment-index=2}

$R^2$ increases When we add a random variable 😦

:::

::: {.fragment fragment-index=3}
However, this is certainly a case where adding a randomly generated variable is *not worth it*. 
:::


:::
::::


::: {.fragment fragment-index=4}
We want to apply the principle of  [Occam's razor](https://www.britannica.com/topic/Occams-razor){target="_blank"}, and choose the simpler model if the increase in $R^2$ is not worth it. 
:::

## Testing gain in prediction: $\Delta R^2$

In hierarchical regression we check whether adding variables to a model yields a significant $R^2$ improvement ($\Delta R^2$). We can use the `anova()` function to compare nested regression models. The *F*-test tells us whether the $R^2$ improvement is significant.

:::: {.columns}
::: {.column width="50%"}

::: {.fragment fragment-index=1}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-115

anova(reg_age, reg_age_fr)
```
<div style="font-size: 23px"> Here, adding `friends` provides a significant improvement in variance explained, $\Delta R^2 = .1, F(1, 97) = 11.47, p = .001$  </div>
:::

:::

::: {.column width="50%"}

::: {.fragment fragment-index=2}
```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

anova(reg_age_fr, reg_age_fr_rand)
```

<div style="font-size: 23px"> Unsurprisingly, adding adding a random variable does not provide a significant improvement in variance explained, $\Delta R^2 = .01, F(1, 96) = 0.81, p = .37$. </div>
:::

:::
::::

<br>


::: {.fragment fragment-index=3}
<div style="font-size: 20px"> **Note**: For reporting, you need to calculate the difference between the  $R^2$ of the two regression models, the $\Delta R^2$, by hand. </div>
:::


## Information Criteria for Model Selection

There also exist a family of statistics used for comparing model based on [information theory](https://en.wikipedia.org/wiki/Information_theory){target="_blank"}. What you will often see are the Akaike information criterion [AIC\; @Akaike_1974] and Bayesian information criterion [BIC\;  @Schwarz_1978]. 

::: {.fragment fragment-index=1}
Both AIC and BIC calculate model fit, and then penalize it based on model complexity (number of estimated [parameters](https://en.wikipedia.org/wiki/Statistical_parameter){target="_blank"}). The logic is that if you are going to add an extra variable to a regression, it should improve fit enough to offset the penalty term. For regression:
:::



:::: {.columns}
::: {.column width="50%"}

::: {.fragment fragment-index=2}

$$AIC = N \times ln(\frac{SS_{\mathrm{resisuals}}}{N}) + 2p$$ 

$N =$ sample size

$p =$ number of model parameters (i.e., intercept, slopes, residual variance)

$SS =$ sum of squares of the residuals 

$ln() =$ stands for the natural log


:::

:::
::: {.column width="50%"}

::: {.fragment fragment-index=2}

$$BIC =  N \times ln(\frac{SS_{\mathrm{resisuals}}}{N}) + p\times ln(N)$$

:::

::: {.fragment fragment-index=3}

The $N \times ln(\frac{SS_{\mathrm{resisuals}}}{N})$ part, which is shared by both formulas, measures model fit. In contrast $2p$ is the AIC penalty, while $p\times ln(N)$ is the BIC penalty. The BIC penalty is stricter.
:::

:::
::::

## The `AIC()` and `BIC()` Functions

We can calculate both AIC and BIC for all our models at once with the `AIC()` and `BIC()` functions. Both for AIC and BIC, the smallest value is the model that fits best according to the criteria. 

:::: {.columns}
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

AIC(reg_age, reg_age_fr, reg_age_fr_rand)
```

The second model fits best.

:::
::: {.column width="50%"}

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

BIC(reg_age,reg_age_fr, reg_age_fr_rand)
```

The second model fits best. 

:::
::::


::: {.fragment fragment-index=1}
In both cases the second model fits best because the addition of the random variable is not enough to offset the penalty for adding an additional variable (ok, that is to be expected, we added a random variable).
:::

::: {.fragment fragment-index=2}
Although, AIC and BIC can disagree (often in my experience!). Usually, AIC likes more complex models, while BIC prefers simpler models. If I have to choose, I tend to trust BIC more.
:::

::: {.fragment fragment-index=3}
Finally, information criteria like the AIC and BIC can be used to compare non-nested models (the data must be the same though), unlike $\Delta R^2$, which can only be used to compare nested models. 
:::

## Caveats With AIC and BIC

There are a couple of caveats to be aware of when using AIC and BIC.

<ol style="font-size: 26px">  


::: {.fragment fragment-index=1}
<li>  Different functions/software will calculate AIC and BIC differently (I was confused for a bit myself before finding [this](https://robjhyndman.com/hyndsight/lm_aic.html){target="_blank"} and [this](https://stackoverflow.com/questions/64531074/why-do-statsmodels-and-r-disagree-on-aic-computation){target="_blank"}). You should not compare AIC and BIC if they come from different functions/software (or, if you need to, be *very* careful).

```{r}
#| eval: true
#| echo: true 
#| code-line-numbers: false
#| classes: code-125

AIC(reg_age)
extractAIC(reg_age)[2]
```

The `AIC()` function adds a $N \times \mathrm{ln}(2\pi) + N$ to the AIC (for math reasons), but other functions may not. So always use the same function/software to calculate AIC and BIC. 


</li>

:::

::: {.fragment fragment-index=2}
<li> AIC and BIC do not tell you *how much better* a model fits over another. So, you may reject a model that practically is quite similar to the model you chose, but you will not know that by looking at AIC and BIC alone. Your theory should also guide the model selection choices that you make. </li>
:::

</ul>

## References 

<div id="refs"> </div>

